---
title: "no-2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries

```{r}
library(tidyverse)  # attaches purrr and readr
library(lubridate)
library(rvest)
library(downloader)
library(R.utils)
library(rlist)
library(here)
library(janitor)
library(scales)
library(ggrepel)
library(ggplot2)
library(here)
library(tidycensus)
library(ggthemes)
library(scales)
library(mapview)
library(here)
library(svMisc)
library(blscrapeR)

# Store census API key
census_api_key("2badea95d03037d0582d8c5ada5195100b1d92f6")

# the function that should be in Tidyverse but isn't
'%!in%' <- function(x,y)!('%in%'(x,y))

options(scipen=999)
options(digits=2)
```

```{r}

#clear environment for when I need it
rm(list = ls())

#why this function isn't in tidyverse, I do not know
'%!in%' <- function(x,y)!('%in%'(x,y))

#function for calculating age, updates to the day; stolen from the internet
calc_age <- function(birthDate, refDate = Sys.Date()) {
    require(lubridate)
    period <- as.period(interval(birthDate, refDate),
                        unit = "year")
    period$year
}

print(Sys.Date())

```


#Import and Clean Data

The data had to be downloaded by month by request of Okaloosa County Clerk. They said larger timeframes would tax their system. This means the data had to be downloaded by month, imported and binded together using a for loop.


```{r}

# list of file names
files <- c(
    "0110", "0210", "0310", "0410", "0510", "0610", "0710", "0810", "0910", "1010", "1110", "1210",
    "0111", "0211", "0311", "0411", "0511", "0611", "0711", "0811", "0911", "1011", "1111", "1211",
    "0112", "0212", "0312", "0412", "0512", "0612", "0712", "0812", "0912", "1012", "1112", "1212",
    "0113", "0213", "0313", "0413", "0513", "0613", "0713", "0813", "0913", "1013", "1113", "1213",
    "0114", "0214", "0314", "0414", "0514", "0614", "0714", "0814", "0914", "1014", "1114", "1214",
    "0115", "0215", "0315", "0415", "0515", "0615", "0715", "0815", "0915", "1015", "1115", "1215",
    "0116", "0216", "0316", "0416", "0516", "0616", "0716", "0816", "0916", "1016", "1116", "1216", 
    "0117", "0217", "0317", "0417", "0517", "0617", "0717", "0817", "0917", "1017", "1117", "1217", 
    "0118", "0218", "0318", "0418", "0518", "0618", "0718", "0818", "0918", "1018", "1118", "1218", 
    "0119", "0219", "0319", "0419", "0519", "0619", "0719", "0819", "0919", "1019", "1119", "1219"
    )

# creating empty dataframe, changes to character were made playing wack-a-mole with binding errors
master_data <- read_csv(here("data", "0115.csv")) %>%
  clean_names() %>%
  filter(arrest_charge == "nothing") %>%
  mutate(actual_speed = as.character(actual_speed),
         blood_alcohol_level = as.character(blood_alcohol_level),
         commercial_vehicle = as.character(commercial_vehicle),
         haul_hazard = as.character(haul_hazard),
         obts_number = as.character(obts_number),
         vehicle_tag_expiration_year = as.character(vehicle_tag_expiration_year),
         prosecutor_charge_count = as.character(prosecutor_charge_count))

# importing, binding to empty data frame 
for (i in files) {
  #writing filepath
  name <- paste0(i, ".csv")
  filepath <- "data/"
  name_plus_filepath <- paste0(filepath,name)
  # importing and cleaning files to temp dataframe
  temp <- read_csv(name_plus_filepath) %>%
    clean_names() 
  temp <- unique(temp)
  # changes to character were made playing wack-a-mole with binding errors
  temp <- temp %>%
  mutate(actual_speed = as.character(actual_speed),
         blood_alcohol_level = as.character(blood_alcohol_level),
         commercial_vehicle = as.character(commercial_vehicle),
         haul_hazard = as.character(haul_hazard),
         obts_number = as.character(obts_number),
         vehicle_tag_expiration_year = as.character(vehicle_tag_expiration_year),
         prosecutor_charge_count = as.character(prosecutor_charge_count),
         zip = as.character(zip),
         bond_hearing_date = as.character(bond_hearing_date),
         arrest_charge_count = as.numeric(arrest_charge_count))
  # binding temp to master_data
  master_data <- master_data %>%
    bind_rows(temp)
  # processing and clearing environment
  rm(temp, name, filepath, name_plus_filepath)
}

rm(i)
rm(files)

# changing columns back to numerics, go lowercase
master_data <- master_data %>%
    mutate_all(tolower) %>%  
    mutate(actual_speed = as.double(actual_speed),
         blood_alcohol_level = as.double(blood_alcohol_level),
         commercial_vehicle = as.double(commercial_vehicle),
         haul_hazard = as.double(haul_hazard),
         obts_number = as.double(obts_number),
         vehicle_tag_expiration_year = as.double(vehicle_tag_expiration_year),
         prosecutor_charge_count = as.double(prosecutor_charge_count),
         case_open_date = mdy(case_open_date),
         case_close_date = mdy(case_close_date),
         disposition_date = mdy(disposition_date),
         date_of_birth = mdy(date_of_birth),
         offense_date = mdy(offense_date),
         arrest_date = mdy(arrest_date),
         arrest_month = month(arrest_date),
         arrest_year = year(arrest_date),
         offense_month = month(offense_date),
         offense_year = year(offense_date),
         close_year = year(case_close_date),
         total_assessed = as.numeric(total_assessed),
         balance = as.numeric(balance),
         total_paid = as.numeric(total_paid),
         obts_number = as.character(obts_number),
         arrest_charge_count = as.numeric(arrest_charge_count),
         arrest_week = week(arrest_date),
         next_day = (arrest_date + 1)
         ) %>%
    filter(offense_year %in% c("2010", "2011", "2012", "2013", "2014", "2015", "2016", "2017", "2018", "2019", "2020")) %>%
    select(-case_type, -disposition_date, -disposition, -defendant_code, -bond_hearing_date, -posted_speed, -actual_speed, -school_status, -commercial_vehicle, -haul_hazard, -d6_issued_date, -d6_issue_status, -d6_released_date)

master_data <- master_data %>%
    unique()

master_data <- master_data %>%
  mutate(case_open_month = substr(case_open_date, start = 1, stop = 7))

charges_by_month <- master_data %>%
  group_by(case_open_month) %>%
  summarise(count = n())

rm(temp)

# There are a few people who have errors in the data that have been manually cleaned:

# In every case, typos in name, birthday or drivers license were tested using a combination of the other two datapoints to prove the third

# Example: donavan jason webb - two different birthdays ("1977-12-16" "1977-02-19") were confirmed errors by # comparing driver's license; a mispelling of his name exists

master_data <- master_data %>%
  mutate(
    date_of_birth = case_when(
      defendant == "donavan jason webb" ~ as_date("1977-02-16"),
      TRUE  ~ date_of_birth
    )
  ) %>%
  mutate(
    defendant = case_when(
      defendant == "donovan jason webb" ~ "donavan jason webb",
      TRUE  ~ defendant
    )
  ) %>%
  #tested with birthday
  mutate(
    drivers_license_number = case_when(
      defendant == "billy ray bottoms" ~ "b352076490060",
      TRUE  ~ drivers_license_number
    )
  ) %>%
  mutate(
  date_of_birth = case_when(
    defendant == "billy ray bottoms" ~ as_date("1949-01-06"),
    TRUE  ~ date_of_birth
    )
  ) %>%
  #tested with birthday
  mutate(
    drivers_license_number = case_when(
      defendant == "timothy roger linville" ~ "l514816561740",
      TRUE  ~ drivers_license_number
    )
  ) %>%
  #tested with driver's license number
  mutate(
  date_of_birth = case_when(
    defendant == "timothy roger linville" ~ as_date("1956-05-14"),
    TRUE  ~ date_of_birth
    )
  ) %>%
  #tested with driver's license number
  mutate(
  date_of_birth = case_when(
    defendant == "timothy roger linville" ~ as_date("1968-01-07"),
    TRUE  ~ date_of_birth
    )
  ) %>%
  mutate(
  date_of_birth = case_when(
    defendant == "daniel lee stetson" ~ as_date("1955-05-31"),
    TRUE  ~ date_of_birth
    )
  ) %>%  
  # used court records to confirm that the date and drivers license were wrong and that she is the other jennifer gail hunt
  mutate(
  date_of_birth = case_when(
    defendant == "jennifer gail hunt" ~ as_date("1979-06-19"),
    TRUE  ~ date_of_birth
    )
  ) %>%
  mutate(
    drivers_license_number = case_when(
      defendant == "jennifer gail hunt" ~ "h530427797190",
      TRUE  ~ drivers_license_number
    )
  ) %>%
  mutate(
  date_of_birth = case_when(
    defendant == "joseph kevin sullivan" ~ as_date("1964-12-23"),
    TRUE  ~ date_of_birth
    )
  ) %>%
  # used court records to confirm right birthdate
    mutate(
  date_of_birth = case_when(
    defendant == "mary moore erlich" ~ as_date("1952-05-13"),
    TRUE  ~ date_of_birth
    )
  ) %>%
  # this person was confirmed the same using both court records and jail inmate records, jail lists him as having been born in 1958	
      mutate(
  date_of_birth = case_when(
    defendant == "phay chaleunsavanh" ~ as_date("1958-02-06"),
    TRUE  ~ date_of_birth
    )
  ) %>%
    mutate(
    drivers_license_number = case_when(
      defendant == "phay chaleunsavanh" ~ "c452660580460",
      TRUE  ~ drivers_license_number
    )
  ) %>%
        mutate(
  date_of_birth = case_when(
    defendant == "robert terry duckworth" ~ as_date("1967-07-24"),
    TRUE  ~ date_of_birth
    )
  ) %>%
          mutate(
  date_of_birth = case_when(
    defendant == "ronald a foster" ~ as_date("1959-04-16"),
    TRUE  ~ date_of_birth
    )
  ) %>%
      mutate(
    drivers_license_number = case_when(
      defendant == "ronald a foster" ~ "f236721591360",
      TRUE  ~ drivers_license_number
    )
  ) %>%
            mutate(
  date_of_birth = case_when(
    defendant == "ronald dean cordero" ~ as_date("1954-05-24"),
    TRUE  ~ date_of_birth
    )
  ) %>%	
  	            mutate(
  date_of_birth = case_when(
    defendant == "russell kippen ferguson" ~ as_date("1948-06-30"),
    TRUE  ~ date_of_birth
    )
  ) %>%	
         mutate(
    drivers_license_number = case_when(
      defendant == "russell kippen ferguson" ~ "f622731482300",
      TRUE  ~ drivers_license_number
    )
  ) %>%
   mutate(
	  date_of_birth = case_when(
    defendant == "terry antonio mcleod" ~ as_date("1968-03-21"),
    TRUE  ~ date_of_birth
    )
  ) %>%	
         mutate(
    drivers_license_number = case_when(
      defendant == "terry antonio mcleod" ~ "m243801681010",
      TRUE  ~ drivers_license_number
    )
  ) %>%
   mutate(
	  date_of_birth = case_when(
    defendant == "william todd bischoff" ~ as_date("1973-01-23"),
    TRUE  ~ date_of_birth
    )
  ) %>%
     mutate(
	  date_of_birth = case_when(
    defendant == "bryan edward may" ~ as_date("1968-01-07"),
    TRUE  ~ date_of_birth
    )
  ) %>%
 # drivers license matches across all 3 names
  mutate(
    defendant = case_when(
      drivers_license_number == "b652004597540" ~ "andrea kizziah barnes",
      TRUE ~ defendant
    )
  ) %>%
  # dob and license match
  mutate(
	  defendant = case_when(
    defendant == "charles edward ferribee" ~ "charles edward ferribee ii",
    TRUE  ~ defendant
    )
  ) %>%
  mutate(
	  defendant = case_when(
    defendant == "dennis carr" ~ "dennis mitchell carr",
    TRUE  ~ defendant
    )
  ) %>%
    mutate(
	  defendant = case_when(
    defendant == "gary earl hunnicutt" ~ "gary earl hunniecutt",
    TRUE  ~ defendant
    )
  ) %>%
  	    mutate(
	  defendant = case_when(
    defendant == "gary earl hunniecut" ~ "gary earl hunniecutt",
    TRUE  ~ defendant
    )
  ) %>%
 	  	    mutate(
	  defendant = case_when(
    defendant == "george hubert spence" ~ "george hubert spence jr",
    TRUE  ~ defendant
    )
  ) %>%
      mutate(
	  defendant = case_when(
    defendant == "gregory a denney" ~ "gregory alan denney",
    TRUE  ~ defendant
    )
  ) %>%
  	    mutate(
	  defendant = case_when(
    defendant == "gregory allan denney" ~ "gregory alan denney",
    TRUE  ~ defendant
    )
  ) %>%
    	    mutate(
	  defendant = case_when(
    defendant == "james william beller" ~ "james william beller s",
    TRUE  ~ defendant
    )
  ) %>%
      	    mutate(
	  defendant = case_when(
    defendant == "jerald warren green" ~ "jerald warren green jr",
    TRUE  ~ defendant
    )
  ) %>%
       	    mutate(
	  defendant = case_when(
    defendant == "jerald warren green" ~ "jerald warren green jr",
    TRUE  ~ defendant
    )
  ) %>%
         	    mutate(
	  defendant = case_when(
    defendant == "jerald warren green" ~ "jerald warren green jr",
    TRUE  ~ defendant
    )
  ) %>%
         	    mutate(
	  defendant = case_when(
    defendant == "jerry stanley groshon" ~ "jerry stanley groshorn",
    TRUE  ~ defendant
    )
  ) %>%
           	    mutate(
	  defendant = case_when(
    defendant == "johnnie s roberts" ~ "johnnie s roberts jr",
    TRUE  ~ defendant
    )
  ) %>%
             	    mutate(
	  defendant = case_when(
    defendant == "joseph keith sullivan" ~ "joseph kevin sullivan",
    TRUE  ~ defendant
    )
  ) %>%
           	    mutate(
	  defendant = case_when(
    defendant == "kenneth ivan schultz" ~ "kenneth ivan shultz iii",
    TRUE  ~ defendant
    )
  ) %>%
             	    mutate(
	  defendant = case_when(
    defendant == "kenneth ivan shoultz iii" ~ "kenneth ivan shultz iii",
    TRUE  ~ defendant
    )
  ) %>%
               	    mutate(
	  defendant = case_when(
    defendant == "kenneth ivan shultz" ~ "kenneth ivan shultz iii",
    TRUE  ~ defendant
    )
  ) %>%
                 	    mutate(
	  defendant = case_when(
    defendant == "lonnie r duvall" ~ "lonnie rodney duvall",
    TRUE  ~ defendant
    )
  ) %>%
                     	    mutate(
	  defendant = case_when(
    defendant == "lowery allen featherstone" ~ "lowery allen featherson",
    TRUE  ~ defendant
    )
  ) %>%
                     	    mutate(
	  defendant = case_when(
    defendant == "mary moore elrich" ~ "mary moore erlich",
    TRUE  ~ defendant
    )
  ) %>%
                       	    mutate(
	  defendant = case_when(
    defendant == "philip field augustine" ~ "philip schoolfield augustine",
    TRUE  ~ defendant
    )
  ) %>%
                         	    mutate(
	  defendant = case_when(
    defendant == "phillip field augustine" ~ "philip schoolfield augustine",
    TRUE  ~ defendant
    )
  ) %>%
                           	    mutate(
	  defendant = case_when(
    defendant == "phillip schoolcraft augustine" ~ "philip schoolfield augustine",
    TRUE  ~ defendant
    )
  ) %>%
   mutate(
  	  defendant = case_when(
    defendant == "phillip schoolfield augustine" ~ "philip schoolfield augustine",
    TRUE  ~ defendant
    )
  ) %>%
     mutate(
  	  defendant = case_when(
    defendant == "richard l abercrombie" ~ "richard leon abercrombie",
    TRUE  ~ defendant
    )
  ) %>%
       mutate(
  	  defendant = case_when(
    defendant == "robert dale taft" ~ "robert dale taff",
    TRUE  ~ defendant
    )
  ) %>%
         mutate(
  	  defendant = case_when(
    defendant == "robert dale taft" ~ "robert dale taff",
    TRUE  ~ defendant
    )
  ) %>%
           mutate(
  	  defendant = case_when(
    defendant == "robin harvey rude" ~ "robin harvey rhude",
    TRUE  ~ defendant
    )
  ) %>%
           mutate(
  	  defendant = case_when(
    defendant == "robin rhude" ~ "robin harvey rhude",
    TRUE  ~ defendant
    )
  ) %>%
           mutate(
  	  defendant = case_when(
    defendant == "ronald w borisenko" ~ "ronald walter borisenko",
    TRUE  ~ defendant
    )
  ) %>%
             mutate(
  	  defendant = case_when(
    defendant == "roy ronald rowell" ~ "roy ronald rowell jr",
    TRUE  ~ defendant
    )
  ) %>%
               mutate(
  	  defendant = case_when(
    defendant == "russell k ferguson" ~ "russell kippen ferguson",
    TRUE  ~ defendant
    )
  ) %>%
               mutate(
  	  defendant = case_when(
    defendant == "sterling lemark parker" ~ "sterling lamark parker",
    TRUE  ~ defendant
    )
  ) %>%
                   mutate(
  	  defendant = case_when(
    defendant == "terry glen long" ~ "terry glenn long",
    TRUE  ~ defendant
    )
  ) %>%
                     mutate(
  	  defendant = case_when(
    defendant == "william curtis howard" ~ "william curtis howard jr",
    TRUE  ~ defendant
    )
  )

#importing scraped data from the Okaloosa County Jail website
jails1 <- read_csv("data/jail-stays-first.csv")
jails2 <- read_csv("data/jail-stays-second.csv")

#binding and removing entries with no name
jail_stays <-  jails1 %>%
  rbind(jails2) %>%
  filter(last != "none") %>%
  mutate(booking_date = mdy(booking_date),
         release_date = mdy(release_date),
         released_in_am = am(release_time),
         days_in_jail = replace_na((release_date - booking_date), 0),
         name = paste0(first, " ", last),
         arrest_week = week(booking_date),
         arrest_year = year(booking_date)
         )

rm(jails1, jails2)

```

# Filtering for Homeless
Fort Walton Beach Police officer confirms that "General Delivery" is an address law enforcement agencies use when they arrest a homeless person. Fort Walton Beach uses "At Large."

```{r}

# creating a dataframe that only includes addresses that match those we know mean they are homeless
general_delivery <- master_data %>%
  filter(address %in% c("general delivery", "genral delivery", "at large", "genreal delivery", "general deliverly", "homeless", "gerneral delivery", "general deliver", "general delviery", "not listed", "general deivery", "grneral delivery", "428 mclaughlin ave", "307 harbor blvd", "183 eglin parkway ne", "117 windham ave", "300 miracle strip", "21 walter martin rd", "123 anywhere st"))


#filter for list of defendants
homeless_names <- general_delivery %>%
  group_by(defendant) %>%
  summarise(charges = n()) %>%
  arrange(desc(charges))

# converting names to vector to feed into for loop
names <- as.vector(homeless_names$defendant)

#creating blank dataframe to bind in the loop
homeless_data <- master_data %>%
  filter(defendant == "nothing")

for (i in names) {
  # filtering for person
  temp <- master_data %>%
    filter(defendant == i)
  # binding to dataframe
  homeless_data <- homeless_data %>%
    bind_rows(temp)
}

rm(temp, i, general_delivery, homeless_names)
```

# Trespass

This frame creates a dataframe that is all trespass charges filed against all homeless people.

```{r}

# creating a dataframe just on trespass
# creating a list of all trespass charges 
all_charges <- master_data %>%
  group_by(arrest_charge, arrest_statute) %>%
  summarise(count = n())
# filtering for all charges with "tres" in the statute
tres_charges <- all_charges %>%
  filter(str_detect(arrest_statute, "tres"))
tres_charges <- tres_charges %>%
  select(arrest_charge) %>%
  unique()
# converting to vector to feed into for loop
trespass <- as.vector(tres_charges$arrest_charge)

# creating blank dataframe
trespass_data <- homeless_data %>%
  filter(arrest_charge == "nothing")

for (i in trespass) {
  # filter for trespass statute
  temp <- homeless_data %>%
    filter(arrest_charge == i)
  # bind to dataframe
  trespass_data <- trespass_data %>%
    bind_rows(temp)
}

trespass_names <- trespass_data %>%
  group_by(defendant) %>%
  count() %>%
  arrange(desc(n))

trespass_names <- trespass_names[1:54, 1:2]

top_50 <- as.vector(trespass_names$defendant)


#THE REST OF THIS CODE LIMITS THE REFERENCE TO TRESPASS DATA TO ONLY THE TOP 50 PLUS TIES IN MOST CHARGED
# creating blank dataframe
trespass_data <- homeless_data %>%
  filter(arrest_charge == "nothing")

for (i in top_50) {
  # filter for trespass statute
  temp <- homeless_data %>%
    filter(defendant == i)
  # bind to dataframe
  trespass_data <- trespass_data %>%
    bind_rows(temp)
}

```



```{r}



homeless_age <- trespass_data %>%
  group_by(defendant, race, gender, date_of_birth) %>%
  summarise() %>%
  mutate(age = calc_age(date_of_birth)) %>%
  select(-date_of_birth) %>%
  arrange(desc(age))


# average is 58
mean(homeless_age$age)

# median age is 58
median(homeless_age$age)
```

```{r}

race <- homeless_age %>%
  group_by(race) %>%
  summarise(count = n()) %>%
  mutate(perc = count/sum(count))

```

```{r}

gender <- homeless_age %>%
  group_by(gender) %>%
  summarise(count = n()) %>%
  mutate(perc = count/sum(count))


```


```{r}

homelesS_charges <- trespass_data %>%
  group_by(defendant) %>%
  count() 


```

```{r}

# all homeless people charged

homeless_person_charges <- trespass_data %>%
  group_by(defendant) %>%
  count()

# average is 32
mean(homeless_person_charges$n)
# median is 28
median(homeless_person_charges$n)

```

```{r}

homeless_person_fees <- trespass_data %>%
  select(case_number, defendant, total_assessed, total_paid) %>%
  unique() %>%
  group_by(defendant) %>%
  summarise(fees_assessed = sum(total_assessed))

homeless_person_paid <- trespass_data %>%
  select(case_number, defendant, total_assessed, total_paid) %>%
  unique() %>%
  group_by(defendant) %>%
  summarise(fees_paid = sum(total_paid))

homeless_person_money <- homeless_person_fees %>%
  inner_join(homeless_person_paid) %>%
  mutate(total_owed = (fees_assessed - fees_paid))

# Average owed per person is $13,263
mean(homeless_person_money$total_owed)
# Median owed per person is 672
median(homeless_person_money$total_owed)

```


```{r}

tres_jail <- trespass_data %>%
  left_join(jail_stays, by = c("defendant" = "name", "arrest_week" = "arrest_week", "arrest_year" = "arrest_year")) %>%
  unique()


tres_jail$days_in_jail[is.na(tres_jail$days_in_jail)] <- 0

#17 days spent in jail on average
mean(tres_jail$days_in_jail)
#9 days spent in jail median
median(tres_jail$days_in_jail)

```

```{r}


#create column that counts missed court charges that resulted from tresspassing
temp19 <- trespass_data %>%
  group_by(case_number, defendant) %>%
  summarise()

temp20 <- homeless_data %>%
  filter(str_detect(arrest_charge, "843.15")) %>%
  group_by(case_number) %>%
  summarise()

temp21 <- temp19 %>%
  inner_join(temp20, by = "case_number") %>%
  mutate(missed_court_charge = "yes")

tres_missed_court <- temp21 %>%
  group_by(defendant) %>%
  summarise(tres_missed_court = n())

(sum(tres_missed_court$tres_missed_court)/54)
```

