---
title: "basics"
author: "Riin Aljas"
date: "2/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggplot2)
library(data.table)
```
###ALL HOMELESS PEOPLE BY MONEY OWED

13 people are double here, because either their race, sex or bday is missing, but that doesn't change their total owed sums. 

If they have a *receipt_date* then it means they're currently in prison and this is the day they were admitted, 
if they have a *release_date*, then that's most current date they were released from prison. 

```{r }

owed_by_person %>% 
  select(-n, all_fees, all_paid, all_balance, receipt_date, release_date)
  data.table()


```
## WHO HAS BEEN CHARGED THE MOST when it comes to dates 

This table shows charges per date, meaning if an officer finds a person with panhandling with marijuana in their pocket, it will count as one as it was same incident. 

```{r }

final_homeless %>% 
  group_by(full_name, offense_date) %>% 
  count() %>% 
  group_by(full_name) %>% 
  count() %>% 
  arrange(desc(n))
  


```


## WHO HAS BEEN CHARGED THE MOST when it comes to charges

This table shows charges per charge, meaning if an officer finds a person with panhandling with marijuana in their pocket, it will count as two as they're separate charges. there isn't much differences compared to the previous data frame 


```{r}
final_homeless %>% 
  group_by(full_name, case_number, initial_charge) %>% 
  count() %>% 
  group_by(full_name) %>% 
  count() %>% 
  arrange(desc(n))
```



```
## ALL PEOPLE
These are all homeless people, with all columns, you can use the search here to see more information per single case or person 
```
```{r cars}

final_homeless %>%
  select(full_name, initial_charge, offense_date, everything()) %>%
  data.table()


```

###BREAKDOWNS BY CHARGES 

```{r}



```

T
