---
title: "PIT analysis"
author: "Sean Mussenden | Howard Center for Investigative Journalism"
date: "9/9/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



``` {r}
#######################
#### Load Packages ####
#######################

# For debugging rm(list=ls())

rm(list=ls())

library(tidyverse)
library(stringr)
library(lubridate)
library(readxl)
library(janitor)
library(weathermetrics)
library(ggplot2)
library(reshape)
library(writexl)
library(dplyr)
library(naniar)

rm(list=ls())

##############################
##### LOAD DATA ##############
##############################

#Point in Time Homeless Counts from HUD 2011-2018

#Eleven_PIT_Counts <- read_xlsx("../../../data/input-data/2007-2018-PIT-Counts-by-CoC.xlsx", sheet = 8, col_names=TRUE)
#Twelve_PIT_Counts <- read_xlsx("../../../data/input-data/2007-2018-PIT-Counts-by-CoC.xlsx", sheet = 7, col_names=TRUE)
#Thirteen_PIT_Counts <- read_xlsx("../../../data/input-data/2007-2018-PIT-Counts-by-CoC.xlsx", sheet = 6, col_names=TRUE)
Fourteen_PIT_Counts <- read_xlsx("../../../data/input-data/2007-2018-PIT-Counts-by-CoC.xlsx", sheet = 5, col_names=TRUE)
#Fifteen_PIT_Counts <- read_xlsx("../../../data/input-data/2007-2018-PIT-Counts-by-CoC.xlsx", sheet = 4, col_names=TRUE)
#Sixteen_PIT_Counts <- read_xlsx("../../../data/input-data/2007-2018-PIT-Counts-by-CoC.xlsx", sheet = 3, col_names=TRUE)
#Seventeen_PIT_Counts <- read_xlsx("../../../data/input-data/2007-2018-PIT-Counts-by-CoC.xlsx", sheet = 2, col_names=TRUE)
Eighteen_PIT_Counts <- read_xlsx("../../../data/input-data/2007-2018-PIT-Counts-by-CoC.xlsx", sheet = 1, col_names=TRUE)

### Load COC clusters analysis from Zillow paper to get coc population for 2017
zillow_coc <- read_csv("../../../data/input-data/CoC_Cluster.csv")
zillow_coc <- clean_names(zillow_coc) %>%
  select(coc_number = co_c_number, everything())



##############################
##### JOIN DATA ##############
##############################

#Join all years together into one data table

all_pit <- Eighteen_PIT_Counts %>%
  #left_join(Seventeen_PIT_Counts, by="CoC Name") %>%
  #left_join(Sixteen_PIT_Counts, by="CoC Name") %>%
  #left_join(Fifteen_PIT_Counts, by="CoC Name") %>%
  left_join(Fourteen_PIT_Counts, by="CoC Name") %>%
  #left_join(Thirteen_PIT_Counts, by="CoC Name") %>%
  #left_join(Twelve_PIT_Counts, by="CoC Name") %>%
  #left_join(Eleven_PIT_Counts, by="CoC Name") %>%
  clean_names() %>%
  select(coc_number = co_c_number_x, matches("overall_homeless")) %>%
  filter(!str_detect(coc_number,"CT-502|CT-506|CT-503|CT-508|CT-512|CT-505|FL-516|FL-503|MA-513|MA-516|MA-520|MA-511|NJ-518|NJ-503|NY-509|NY-502|NY-505|NY-517|	NY-508|PA-507|PA-509|PA-602|PA-601|TX-703|TX-607|IN-500|IN-502|LA-504|LA-509|MA-518|MA-516|ME-502|ME-500|AR-512|AR-503|LA-508|LA-509|WA-507|WA-501"))%>%
  filter(!is.na(coc_number)) %>%
  filter(!str_detect(coc_number, "territory")) %>%
  filter(!str_detect(coc_number, "AR-512")) %>%
  na.omit() %>%
  left_join(zillow_coc, by="coc_number") %>%
  select(coc_number, coc_name = co_c_name, overall_homeless_2014, overall_homeless_2018, total_population_2017 = co_c_total_population) %>%
  na.omit() %>%
  mutate_at(vars(matches("overall_homeless")), as.numeric) %>%
  mutate(homeless_rate_2014 = overall_homeless_2014/total_population_2017*100,
         homeless_rate_2018 = overall_homeless_2018/total_population_2017*100,
         overall_change_2014_2018 = overall_homeless_2018 -  overall_homeless_2014,
         overall_pct_change_2014_2018 = (overall_homeless_2018 -  overall_homeless_2014)/overall_homeless_2014,
         overall_pct_pt_change_2014_2018 = homeless_rate_2018 - homeless_rate_2014
         )


```


## Zillow Analysis ##

### Fact: Highest rates of homelessness by COC

In 2017, Mendocino County's COC had the nation's highest rate of homelessness, at 2 percent, followed by Washington, D.C. (1.2 percent).  Five of the top 10 were in California (Mendocino County, Santa Cruz area, San Francisco, Monterey area, Imperial County).  Other areas with high rates: Boston, New York City, Atlanta and the Florida Keys (Monroe County).  

### Supporting code and output
``` {r}

zillow_coc %>%
  select(coc_number, co_c_name, estimated_homeless_rate_percent) %>%
  arrange(desc(estimated_homeless_rate_percent))

```

### Fact: Highest rates of homelessness
Washington, D.C., followed by Hawaii, Oregon and California had highest average rates

### Supporting code and output

``` {r}

zillow_coc %>%
  separate(coc_number, sep="-", into= c("state", "number")) %>%
  select(state, co_c_name, estimated_homeless_rate_percent) %>%
  group_by(state) %>%
  summarise(median_homeless_rate = median(estimated_homeless_rate_percent)) %>%
  arrange(desc(median_homeless_rate))
```


### ALL PIT

``` {r}
  all_pit %>%
    arrange(overall_pct_change_2014_2018) %>%
    select(coc_number, coc_name, overall_pct_change_2014_2018, everything())

  all_pit %>%
    arrange(desc(overall_change_2014_2018)) %>%
    select(coc_number, coc_name, overall_change_2014_2018, everything() )

```

##############################
##### FINDING TRENDS IN HOMELESS POP ##############
##############################

##Look at: 2007-2018, 2011-2018, 2014-2018 for overall, sheltered, unsheltered

PIT_counts_most_coc <- eighteen_to_seven_pit[-c(400:4495),] %>% 
  clean_names() %>%
  select(coc_number = co_c_number_x, matches("overall_homeless")) %>%
  mutate(state_name = str_extract_all(coc_number, "\\D\\D")) %>%
  filter(state_name != "CA",
         state_name != "MA",
         state_name != "NY") %>%
  naniar::replace_with_na_all(condition = ~.x == ".") %>%
  mutate_at(vars(matches("overall")), as.numeric) %>%
  mutate(seven_to_eighteen_change = 100*((`overall_homeless_2018` - `overall_homeless_2007`)/`overall_homeless_2007`))

ggplot(data = PIT_counts_most_coc, aes(x=coc_number, y=seven_to_eighteen_change)) +
  geom_point() +
  labs(y = "Percent Change",
       title = "2007 to 2018 Percent Change in Overall Homelessness") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  facet_wrap(~state_name, ncol = 3, scales = "free_x")

ggsave(filename = "PIT_counts_most_coc.png", 
       device = "png", path = "./graphs",
       width = 15, height = 45, units = "in")

PIT_counts_big_coc <- eighteen_to_seven_pit[-c(400:4495),] %>% 
  clean_names() %>%
  select(coc_number = co_c_number_x, matches("overall_homeless")) %>%
  mutate(state_name = str_extract_all(coc_number, "\\D\\D")) %>%
  filter(state_name == "CA" | state_name == "MA" | state_name == "NY") %>%
  naniar::replace_with_na_all(condition = ~.x == ".") %>%
  mutate_at(vars(matches("overall")), as.numeric) %>%
  mutate(seven_to_eighteen_change = 100*((`overall_homeless_2018` - `overall_homeless_2007`)/`overall_homeless_2007`))

ggplot(data = PIT_counts_big_coc, aes(x=coc_number, y=seven_to_eighteen_change)) +
  geom_point() +
  labs(y = "Percent Change",
       title = "NY MA & CA 2007 to 2018 Percent Change in Overall Homelessness") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  facet_wrap(~state_name, ncol = 1, scales = "free_x")

ggsave(filename = "PIT_counts_big_coc.png", 
       device = "png", path = "./graphs",
       width = 15, height = 30, units = "in")