---
title: "NYC 311 Calls"
author: "Theresa Diffendal"
date: "1/27/2020"
output: 
  html_document:
      toc: true
      toc_depth: 3
      toc_float: true
    md_document:
      variant: markdown_github
      toc: true
      toc_depth: 3
---

An analysis of New YOrk City's 311 call data from 2010-present. Data found on New York's [open data site](https://data.cityofnewyork.us/Social-Services/311-Service-Requests-from-2010-to-Present/erm2-nwe9/data).

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

###Libraries

#install.packages("readr")
#install.packages("bit64")

library(tidyverse)
library(janitor)
library(lubridate)
library(ggplot2)
library(readxl)
library(dplyr)
library(tidyr)
library(data.table)
library(readr)
#library(bit64)
```

## Cleaning Work
### Column Names
```{r}
##Load the csv in chunks, clean names, maniuplate date_time columns into separate year, month, and time columns

ccolumns <- list("Unique Key", "Created Date", "Closed Date", "Agency", "Agency Name", "Complaint Type", "Descriptor", "Location Type", "Incident Zip", "Incident Address", "Street Name", "Cross Street 1", "Cross Street 2", "Intersection Street 1", "Intersection Street 2", "Address Type", "City", "Landmark", "Facility Type", "Status", "Due Date", "Resolution Description", "Resolution Action Updated Date", "Community Board", "BBL", "Borough", "X Coordinate (State Plane)", "Y Coordinate (State Plane)", "Open Data Channel Type", "Park Facility Name", "Park Borough", "Vehicle Type", "Taxi Company Borough", "Taxi Pick Up Location", "Bridge Highway Name", "Bridge Highway Direction", "Road Ramp", "Bridge Highway Segment", "Latitude", "Longitude", "Location")

#coltypess <- c("double", "date", "date", "text", "text", "text", "text", "text", "double", "text", "text", "text", "text", "text", "text", "text", "text", "text", "text", "text", "date", "text", "date", "text", "double", "text", "double", "double", "text", "text", "text", "text", "text", "text", "text", "text", "text", "text", "double", "double", "double")

#col_datetime(format = "%m/%d/%Y %I:%M:%S %p")
```


### Read in CSV 1 million lines at a time, save as RDS

```{r}
#########################
##### First Million #####
#########################

sskip = 1
nn_max = 0

while(nn_max < 1e6) {  
  #22.4e6 is the total number of rows in the dataset
  
  nn_max = nn_max + 100000
  
  nyc_calls_01 <- read_csv("~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC_311_Service_Requests_from_2010_to_Present.csv", col_names = as.character(ccolumns), col_types = cols(.default = col_character()), n_max = nn_max, skip = sskip) %>%
    clean_names() %>%
    separate(created_date, c("dates", "hour", "AM/PM"), " ") %>%
    separate("dates", c("month", "day", "year"), remove=FALSE) %>%
    unite("time", "hour":"AM/PM", sep = " ")

  nyc_calls_01$dates <- as.Date(nyc_calls_01$dates, format = "%m/%d/%Y")
  
  sskip = sskip + 100000
}

#save as Rds(for faster processing)
saveRDS(nyc_calls_01, "~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC-311-Service-Requests_01.Rds")




#########################
######Second Million#####
#########################

sskip = 1e6 +1
nn_max = 0

while(nn_max < 1e6) {  
  #22.4e6 is the total number of rows in the dataset
  
  nn_max = nn_max + 100000
  
  nyc_calls_02 <- read_csv("~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC_311_Service_Requests_from_2010_to_Present.csv", col_names = as.character(ccolumns), col_types = cols(.default = col_character()), n_max = nn_max, skip = sskip) %>%
    clean_names() %>%
    separate(created_date, c("dates", "hour", "AM/PM"), " ") %>%
    separate("dates", c("month", "day", "year"), remove=FALSE) %>%
    unite("time", "hour":"AM/PM", sep = " ") 
  
  nyc_calls_02$dates <- as.Date(nyc_calls_02$dates, format = "%m/%d/%Y")
  
  sskip = sskip + 100000
}

#save as Rds(for faster processing)
saveRDS(nyc_calls_02, "~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC-311-Service-Requests_02.Rds")



########################
#####Third Million######
########################

sskip = 2e6 +1
nn_max = 0

while(nn_max < 1e6) {  
  #22.4e6 is the total number of rows in the dataset
  
  nn_max = nn_max + 100000
  
  nyc_calls_03 <- read_csv("~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC_311_Service_Requests_from_2010_to_Present.csv", col_names = as.character(ccolumns), col_types = cols(.default = col_character()), n_max = nn_max, skip = sskip) %>%
    clean_names() %>%
    separate(created_date, c("dates", "hour", "AM/PM"), " ") %>%
    separate("dates", c("month", "day", "year"), remove=FALSE) %>%
    unite("time", "hour":"AM/PM", sep = " ") 

  nyc_calls_03$dates <- as.Date(nyc_calls_03$dates, format = "%m/%d/%Y")
  
  sskip = sskip + 100000
}

#save as Rds(for faster processing)
saveRDS(nyc_calls_03, "~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC-311-Service-Requests_03.Rds")



#########################
#### Fourth Million #####
#########################

sskip = 3e6 +1
nn_max = 0

while(nn_max < 1e6) {  
  #22.4e6 is the total number of rows in the dataset
  
  nn_max = nn_max + 100000
  
  nyc_calls_04 <- read_csv("~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC_311_Service_Requests_from_2010_to_Present.csv", col_names = as.character(ccolumns), col_types = cols(.default = col_character()), n_max = nn_max, skip = sskip) %>%
    clean_names() %>%
    separate(created_date, c("dates", "hour", "AM/PM"), " ") %>%
    separate("dates", c("month", "day", "year"), remove=FALSE) %>%
    unite("time", "hour":"AM/PM", sep = " ")

  nyc_calls_04$dates <- as.Date(nyc_calls_04$dates, format = "%m/%d/%Y")
  
  sskip = sskip + 100000
}

#save as Rds(for faster processing)
saveRDS(nyc_calls_04, "~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC-311-Service-Requests_04.Rds")



#########################
##### Fifth Million #####
#########################

sskip = 4e6 +1
nn_max = 0

while(nn_max < 1e6) {  
  #22.4e6 is the total number of rows in the dataset
  
  nn_max = nn_max + 100000
  
  nyc_calls_05 <- read_csv("~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC_311_Service_Requests_from_2010_to_Present.csv", col_names = as.character(ccolumns), col_types = cols(.default = col_character()), n_max = nn_max, skip = sskip) %>%
    clean_names() %>%
    separate(created_date, c("dates", "hour", "AM/PM"), " ") %>%
    separate("dates", c("month", "day", "year"), remove=FALSE) %>%
    unite("time", "hour":"AM/PM", sep = " ")

  nyc_calls_05$dates <- as.Date(nyc_calls_05$dates, format = "%m/%d/%Y")
  
  sskip = sskip + 100000
}

#save as Rds(for faster processing)
saveRDS(nyc_calls_05, "~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC-311-Service-Requests_05.Rds")


#########################
####### 6 Million #######
#########################

sskip = 5e6 +1
nn_max = 0

while(nn_max < 1e6) {  
  #22.4e6 is the total number of rows in the dataset
  
  nn_max = nn_max + 100000
  
  nyc_calls_06 <- read_csv("~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC_311_Service_Requests_from_2010_to_Present.csv", col_names = as.character(ccolumns), col_types = cols(.default = col_character()), n_max = nn_max, skip = sskip) %>%
    clean_names() %>%
    separate(created_date, c("dates", "hour", "AM/PM"), " ") %>%
    separate("dates", c("month", "day", "year"), remove=FALSE) %>%
    unite("time", "hour":"AM/PM", sep = " ")

  nyc_calls_06$dates <- as.Date(nyc_calls_06$dates, format = "%m/%d/%Y")
  
  sskip = sskip + 100000
}

#save as Rds(for faster processing)
saveRDS(nyc_calls_06, "~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC-311-Service-Requests_06.Rds")




#########################
####### 7 Million #######
#########################

sskip = 6e6 +1
nn_max = 0

while(nn_max < 1e6) {  
  #22.4e6 is the total number of rows in the dataset
  
  nn_max = nn_max + 100000
  
  nyc_calls_07 <- read_csv("~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC_311_Service_Requests_from_2010_to_Present.csv", col_names = as.character(ccolumns), col_types = cols(.default = col_character()), n_max = nn_max, skip = sskip) %>%
    clean_names() %>%
    separate(created_date, c("dates", "hour", "AM/PM"), " ") %>%
    separate("dates", c("month", "day", "year"), remove=FALSE) %>%
    unite("time", "hour":"AM/PM", sep = " ")

  nyc_calls_07$dates <- as.Date(nyc_calls_07$dates, format = "%m/%d/%Y")
  
  sskip = sskip + 100000
}

#save as Rds(for faster processing)
saveRDS(nyc_calls_07, "~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC-311-Service-Requests_07.Rds")




#########################
####### 8 Million #######
#########################

sskip = 7e6 +1
nn_max = 0

while(nn_max < 1e6) {  
  #22.4e6 is the total number of rows in the dataset
  
  nn_max = nn_max + 100000
  
  nyc_calls_08 <- read_csv("~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC_311_Service_Requests_from_2010_to_Present.csv", col_names = as.character(ccolumns), col_types = cols(.default = col_character()), n_max = nn_max, skip = sskip) %>%
    clean_names() %>%
    separate(created_date, c("dates", "hour", "AM/PM"), " ") %>%
    separate("dates", c("month", "day", "year"), remove=FALSE) %>%
    unite("time", "hour":"AM/PM", sep = " ")

  nyc_calls_08$dates <- as.Date(nyc_calls_08$dates, format = "%m/%d/%Y")
  
  sskip = sskip + 100000
}

#save as Rds(for faster processing)
saveRDS(nyc_calls_08, "~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC-311-Service-Requests_08.Rds")




#########################
####### 9 Million #######
#########################

sskip = 8e6 +1
nn_max = 0

while(nn_max < 1e6) {  
  #22.4e6 is the total number of rows in the dataset
  
  nn_max = nn_max + 100000
  
  nyc_calls_09 <- read_csv("~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC_311_Service_Requests_from_2010_to_Present.csv", col_names = as.character(ccolumns), col_types = cols(.default = col_character()), n_max = nn_max, skip = sskip) %>%
    clean_names() %>%
    separate(created_date, c("dates", "hour", "AM/PM"), " ") %>%
    separate("dates", c("month", "day", "year"), remove=FALSE) %>%
    unite("time", "hour":"AM/PM", sep = " ")

  nyc_calls_09$dates <- as.Date(nyc_calls_09$dates, format = "%m/%d/%Y")
  
  sskip = sskip + 100000
}

#save as Rds(for faster processing)
saveRDS(nyc_calls_09, "~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC-311-Service-Requests_09.Rds")



#########################
####### 10 Million ######
#########################

sskip = 9e6 +1
nn_max = 0

while(nn_max < 1e6) {  
  #22.4e6 is the total number of rows in the dataset
  
  nn_max = nn_max + 100000
  
  nyc_calls_10 <- read_csv("~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC_311_Service_Requests_from_2010_to_Present.csv", col_names = as.character(ccolumns), col_types = cols(.default = col_character()), n_max = nn_max, skip = sskip) %>%
    clean_names() %>%
    separate(created_date, c("dates", "hour", "AM/PM"), " ") %>%
    separate("dates", c("month", "day", "year"), remove=FALSE) %>%
    unite("time", "hour":"AM/PM", sep = " ")

  nyc_calls_10$dates <- as.Date(nyc_calls_10$dates, format = "%m/%d/%Y")
  
  sskip = sskip + 100000
}

#save as Rds(for faster processing)
saveRDS(nyc_calls_10, "~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC-311-Service-Requests_10.Rds")



#########################
###### 11 Million #######
#########################

sskip = 1e7 +1
nn_max = 0

while(nn_max < 1e6) {  
  #22.4e6 is the total number of rows in the dataset
  
  nn_max = nn_max + 100000
  
  nyc_calls_11 <- read_csv("~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC_311_Service_Requests_from_2010_to_Present.csv", col_names = as.character(ccolumns), col_types = cols(.default = col_character()), n_max = nn_max, skip = sskip) %>%
    clean_names() %>%
    separate(created_date, c("dates", "hour", "AM/PM"), " ") %>%
    separate("dates", c("month", "day", "year"), remove=FALSE) %>%
    unite("time", "hour":"AM/PM", sep = " ")

  nyc_calls_11$dates <- as.Date(nyc_calls_11$dates, format = "%m/%d/%Y")
  
  sskip = sskip + 100000
}

#save as Rds(for faster processing)
saveRDS(nyc_calls_11, "~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC-311-Service-Requests_11.Rds")





#########################
###### 12 Million #######
#########################

sskip = 11e6 +1
nn_max = 0

while(nn_max < 1e6) {  
  #22.4e6 is the total number of rows in the dataset
  
  nn_max = nn_max + 100000
  
  nyc_calls_12 <- read_csv("~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC_311_Service_Requests_from_2010_to_Present.csv", col_names = as.character(ccolumns), col_types = cols(.default = col_character()), n_max = nn_max, skip = sskip) %>%
    clean_names() %>%
    separate(created_date, c("dates", "hour", "AM/PM"), " ") %>%
    separate("dates", c("month", "day", "year"), remove=FALSE) %>%
    unite("time", "hour":"AM/PM", sep = " ")

  nyc_calls_12$dates <- as.Date(nyc_calls_12$dates, format = "%m/%d/%Y")
  
  sskip = sskip + 100000
}

#save as Rds(for faster processing)
saveRDS(nyc_calls_12, "~/Documents/GitHub/homelessness-in-the-us/new-york-311//NYC-311-Service-Requests_12.Rds")






#########################
###### 13 Million #######
#########################

sskip = 12e6 +1
nn_max = 0

while(nn_max < 1e6) {  
  #22.4e6 is the total number of rows in the dataset
  
  nn_max = nn_max + 100000
  
  nyc_calls_13 <- read_csv("~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC_311_Service_Requests_from_2010_to_Present.csv", col_names = as.character(ccolumns), col_types = cols(.default = col_character()), n_max = nn_max, skip = sskip) %>%
    clean_names() %>%
    separate(created_date, c("dates", "hour", "AM/PM"), " ") %>%
    separate("dates", c("month", "day", "year"), remove=FALSE) %>%
    unite("time", "hour":"AM/PM", sep = " ")

  nyc_calls_13$dates <- as.Date(nyc_calls_13$dates, format = "%m/%d/%Y")
  
  sskip = sskip + 100000
}

#save as Rds(for faster processing)
saveRDS(nyc_calls_13, "~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC-311-Service-Requests_13.Rds")




#########################
##### 14 Million #####
#########################

sskip = 13e6 +1
nn_max = 0

while(nn_max < 1e6) {  
  #22.4e6 is the total number of rows in the dataset
  
  nn_max = nn_max + 100000
  
  nyc_calls_14 <- read_csv("~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC_311_Service_Requests_from_2010_to_Present.csv", col_names = as.character(ccolumns), col_types = cols(.default = col_character()), n_max = nn_max, skip = sskip) %>%
    clean_names() %>%
    separate(created_date, c("dates", "hour", "AM/PM"), " ") %>%
    separate("dates", c("month", "day", "year"), remove=FALSE) %>%
    unite("time", "hour":"AM/PM", sep = " ")

  nyc_calls_14$dates <- as.Date(nyc_calls_14$dates, format = "%m/%d/%Y")
  
  sskip = sskip + 100000
}

#save as Rds(for faster processing)
saveRDS(nyc_calls_14, "~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC-311-Service-Requests_14.Rds")




#########################
##### 15 Million #####
#########################

sskip = 14e6 +1
nn_max = 0

while(nn_max < 1e6) {  
  #22.4e6 is the total number of rows in the dataset
  
  nn_max = nn_max + 100000
  
  nyc_calls_15 <- read_csv("~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC_311_Service_Requests_from_2010_to_Present.csv", col_names = as.character(ccolumns), col_types = cols(.default = col_character()), n_max = nn_max, skip = sskip) %>%
    clean_names() %>%
    separate(created_date, c("dates", "hour", "AM/PM"), " ") %>%
    separate("dates", c("month", "day", "year"), remove=FALSE) %>%
    unite("time", "hour":"AM/PM", sep = " ")

  nyc_calls_15$dates <- as.Date(nyc_calls_15$dates, format = "%m/%d/%Y")
  
  sskip = sskip + 100000
}

#save as Rds(for faster processing)
saveRDS(nyc_calls_15, "~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC-311-Service-Requests_15.Rds")

#########################
###### 16 Million #######
#########################

sskip = 15e6 +1
nn_max = 0

while(nn_max < 1e6) {  
  #22.4e6 is the total number of rows in the dataset
  
  nn_max = nn_max + 100000
  
  nyc_calls_16 <- read_csv("~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC_311_Service_Requests_from_2010_to_Present.csv", col_names = as.character(ccolumns), col_types = cols(.default = col_character()), n_max = nn_max, skip = sskip) %>%
    clean_names() %>%
    separate(created_date, c("dates", "hour", "AM/PM"), " ") %>%
    separate("dates", c("month", "day", "year"), remove=FALSE) %>%
    unite("time", "hour":"AM/PM", sep = " ")

  nyc_calls_16$dates <- as.Date(nyc_calls_16$dates, format = "%m/%d/%Y")
  
  sskip = sskip + 100000
}

#save as Rds(for faster processing)
saveRDS(nyc_calls_16, "~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC-311-Service-Requests_16.Rds")



#########################
###### 17 Million #######
#########################

sskip = 16e6 +1
nn_max = 0

while(nn_max < 1e6) {  
  #22.4e6 is the total number of rows in the dataset
  
  nn_max = nn_max + 100000
  
  nyc_calls_17 <- read_csv("~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC_311_Service_Requests_from_2010_to_Present.csv", col_names = as.character(ccolumns), col_types = cols(.default = col_character()), n_max = nn_max, skip = sskip) %>%
    clean_names() %>%
    separate(created_date, c("dates", "hour", "AM/PM"), " ") %>%
    separate("dates", c("month", "day", "year"), remove=FALSE) %>%
    unite("time", "hour":"AM/PM", sep = " ")

  nyc_calls_17$dates <- as.Date(nyc_calls_17$dates, format = "%m/%d/%Y")
  
  sskip = sskip + 100000
}

#save as Rds(for faster processing)
saveRDS(nyc_calls_17, "~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC-311-Service-Requests_17.Rds")



#########################
##### 18 Million #####
#########################

sskip = 17e6 +1
nn_max = 0

while(nn_max < 1e6) {  
  #22.4e6 is the total number of rows in the dataset
  
  nn_max = nn_max + 100000
  
  nyc_calls_18 <- read_csv("~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC_311_Service_Requests_from_2010_to_Present.csv", col_names = as.character(ccolumns), col_types = cols(.default = col_character()), n_max = nn_max, skip = sskip) %>%
    clean_names() %>%
    separate(created_date, c("dates", "hour", "AM/PM"), " ") %>%
    separate("dates", c("month", "day", "year"), remove=FALSE) %>%
    unite("time", "hour":"AM/PM", sep = " ")

  nyc_calls_18$dates <- as.Date(nyc_calls_18$dates, format = "%m/%d/%Y")
  
  sskip = sskip + 100000
}

#save as Rds(for faster processing)
saveRDS(nyc_calls_18, "~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC-311-Service-Requests_18.Rds")



#########################
###### 19 Million #######
#########################

sskip = 18e6 +1
nn_max = 0

while(nn_max < 1e6) {  
  #22.4e6 is the total number of rows in the dataset
  
  nn_max = nn_max + 100000
  
  nyc_calls_19 <- read_csv("~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC_311_Service_Requests_from_2010_to_Present.csv", col_names = as.character(ccolumns), col_types = cols(.default = col_character()), n_max = nn_max, skip = sskip) %>%
    clean_names() %>%
    separate(created_date, c("dates", "hour", "AM/PM"), " ") %>%
    separate("dates", c("month", "day", "year"), remove=FALSE) %>%
    unite("time", "hour":"AM/PM", sep = " ")

  nyc_calls_19$dates <- as.Date(nyc_calls_19$dates, format = "%m/%d/%Y")
  
  sskip = sskip + 100000
}

#save as Rds(for faster processing)
saveRDS(nyc_calls_19, "~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC-311-Service-Requests_19.Rds")



 
#########################
###### 20 Million #######
#########################

sskip = 19e6 +1
nn_max = 0

while(nn_max < 1e6) {  
  #22.4e6 is the total number of rows in the dataset
  
  nn_max = nn_max + 100000
  
  nyc_calls_20 <- read_csv("~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC_311_Service_Requests_from_2010_to_Present.csv", col_names = as.character(ccolumns), col_types = cols(.default = col_character()), n_max = nn_max, skip = sskip) %>%
    clean_names() %>%
    separate(created_date, c("dates", "hour", "AM/PM"), " ") %>%
    separate("dates", c("month", "day", "year"), remove=FALSE) %>%
    unite("time", "hour":"AM/PM", sep = " ")

  nyc_calls_20$dates <- as.Date(nyc_calls_20$dates, format = "%m/%d/%Y")
  
  sskip = sskip + 100000
}

#save as Rds(for faster processing)
saveRDS(nyc_calls_20, "~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC-311-Service-Requests_20.Rds")



#########################
##### 21 Million #####
#########################

sskip = 2e7 +1
nn_max = 0

while(nn_max < 1e6) {  
  #22.4e6 is the total number of rows in the dataset
  
  nn_max = nn_max + 100000
  
  nyc_calls_21 <- read_csv("~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC_311_Service_Requests_from_2010_to_Present.csv", col_names = as.character(ccolumns), col_types = cols(.default = col_character()), n_max = nn_max, skip = sskip) %>%
    clean_names() %>%
    separate(created_date, c("dates", "hour", "AM/PM"), " ") %>%
    separate("dates", c("month", "day", "year"), remove=FALSE) %>%
    unite("time", "hour":"AM/PM", sep = " ")

  nyc_calls_21$dates <- as.Date(nyc_calls_21$dates, format = "%m/%d/%Y")
  
  sskip = sskip + 100000
}

#save as Rds(for faster processing)
saveRDS(nyc_calls_21, "~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC-311-Service-Requests_21.Rds")




#########################
###### 22 Million #######
#########################

sskip = 21e6 +1
nn_max = 0

while(nn_max < 1e6) {  
  #22.4e6 is the total number of rows in the dataset
  
  nn_max = nn_max + 100000
  
  nyc_calls_22 <- read_csv("~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC_311_Service_Requests_from_2010_to_Present.csv", col_names = as.character(ccolumns), col_types = cols(.default = col_character()), n_max = nn_max, skip = sskip) %>%
    clean_names() %>%
    separate(created_date, c("dates", "hour", "AM/PM"), " ") %>%
    separate("dates", c("month", "day", "year"), remove=FALSE) %>%
    unite("time", "hour":"AM/PM", sep = " ")

  nyc_calls_22$dates <- as.Date(nyc_calls_22$dates, format = "%m/%d/%Y")
  
  sskip = sskip + 100000
}

#save as Rds(for faster processing)
saveRDS(nyc_calls_22, "~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC-311-Service-Requests_22.Rds")



#########################
###### 23 Million #######
#########################

sskip = 22e6 +1
nn_max = 0

while(nn_max < 1e6) {  
  #22.4e6 is the total number of rows in the dataset
  
  nn_max = nn_max + 100000
  
  nyc_calls_23 <- read_csv("~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC_311_Service_Requests_from_2010_to_Present.csv", col_names = as.character(ccolumns), col_types = cols(.default = col_character()), n_max = nn_max, skip = sskip) %>%
    clean_names() %>%
    separate(created_date, c("dates", "hour", "AM/PM"), " ") %>%
    separate("dates", c("month", "day", "year"), remove=FALSE) %>%
    unite("time", "hour":"AM/PM", sep = " ")

  nyc_calls_23$dates <- as.Date(nyc_calls_23$dates, format = "%m/%d/%Y")
  
  sskip = sskip + 100000
}

#save as Rds(for faster processing)
saveRDS(nyc_calls_23, "~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC-311-Service-Requests_23.Rds")

```

### Create all NYC calls dataset
```{r}

# Read in the multiple RDS files and filter by date

calls_01 <- readRDS("~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC-311-Service-Requests_01.Rds")
calls_02 <- readRDS("~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC-311-Service-Requests_02.Rds")
calls_03 <- readRDS("~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC-311-Service-Requests_03.Rds")
calls_04 <- readRDS("~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC-311-Service-Requests_04.Rds")
calls_05 <- readRDS("~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC-311-Service-Requests_05.Rds")
calls_06 <- readRDS("~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC-311-Service-Requests_06.Rds")
calls_07 <- readRDS("~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC-311-Service-Requests_07.Rds")
calls_08 <- readRDS("~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC-311-Service-Requests_08.Rds")
calls_09 <- readRDS("~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC-311-Service-Requests_09.Rds")
calls_10 <- readRDS("~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC-311-Service-Requests_10.Rds")
calls_11 <- readRDS("~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC-311-Service-Requests_11.Rds")
calls_12 <- readRDS("~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC-311-Service-Requests_12.Rds")
calls_13 <- readRDS("~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC-311-Service-Requests_13.Rds")
calls_14 <- readRDS("~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC-311-Service-Requests_14.Rds")
calls_15 <- readRDS("~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC-311-Service-Requests_15.Rds")
calls_16 <- readRDS("~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC-311-Service-Requests_16.Rds")
calls_17 <- readRDS("~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC-311-Service-Requests_17.Rds")
calls_18 <- readRDS("~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC-311-Service-Requests_18.Rds")
calls_19 <- readRDS("~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC-311-Service-Requests_19.Rds")
calls_20 <- readRDS("~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC-311-Service-Requests_20.Rds")
calls_21 <- readRDS("~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC-311-Service-Requests_21.Rds")
calls_22 <- readRDS("~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC-311-Service-Requests_22.Rds")
calls_23 <- readRDS("~/Documents/GitHub/homelessness-in-the-us/new-york-311/NYC-311-Service-Requests_23.Rds")
  

#create object with file names and full paths
#f <- file.path("~/Documents/GitHub/homelessness-in-the-us/new-york-311", c("NYC-311-Service-Requests_01.Rds", "NYC-311-Service-Requests_02.Rds", "NYC-311-Service-Requests_03.Rds", "NYC-311-Service-Requests_04.Rds", "NYC-311-Service-Requests_05.Rds", "NYC-311-Service-Requests_06.Rds", "NYC-311-Service-Requests_07.Rds", "NYC-311-Service-Requests_08.Rds", "NYC-311-Service-Requests_09.Rds", "NYC-311-Service-Requests_10.Rds", "NYC-311-Service-Requests_11.Rds", "NYC-311-Service-Requests_12.Rds", "NYC-311-Service-Requests_13.Rds", "NYC-311-Service-Requests_14.Rds", "NYC-311-Service-Requests_15.Rds", "NYC-311-Service-Requests_16.Rds", "NYC-311-Service-Requests_17.Rds", "NYC-311-Service-Requests_18.Rds", "NYC-311-Service-Requests_19.Rds", "NYC-311-Service-Requests_20.Rds", "NYC-311-Service-Requests_21.Rds", "NYC-311-Service-Requests_22.Rds", "NYC-311-Service-Requests_23.Rds"))


#bind RDS files to an all in one
all_nyc_calls <- rbind(calls_01, calls_02, calls_03, calls_04, calls_05, calls_06, calls_07, calls_08, calls_09, calls_10, calls_11, calls_12, calls_13, calls_14, calls_15, calls_16, calls_17, calls_18, calls_19, calls_20, calls_21, calls_22) 

saveRDS(all_nyc_calls, "~/Documents/GitHub/homelessness-in-the-us/new-york-311/ALL_NYC-311-Service-Requests.Rds")
```

### Create RDS sheet with each year, filtered from all NYC calls
```{r}
#Filter all calls for each year and save RDS

#2010
calls_2010 <- all_nyc_calls %>%
  filter(year == 2010) 
calls_2010[order(calls_2010$month, calls_2010$day, calls_2010$time),]
saveRDS(calls_2010, "~/Documents/GitHub/homelessness-in-the-us/new-york-311/2010-NYC-Service-Requests.Rds")

#2011
calls_2011 <- all_nyc_calls %>%
  filter(year == 2011) 
calls_2011[order(calls_2011$month, calls_2011$day, calls_2011$time),]
saveRDS(calls_2011, "~/Documents/GitHub/homelessness-in-the-us/new-york-311/2011-NYC-Service-Requests.Rds")

#2012 
calls_2012 <- all_nyc_calls %>%
  filter(year == 2012) 
calls_2012[order(calls_2012$month, calls_2012$day, calls_2012$time),]
saveRDS(calls_2012, "~/Documents/GitHub/homelessness-in-the-us/new-york-311/2012-NYC-Service-Requests.Rds")

#2013
calls_2013 <- all_nyc_calls %>%
  filter(year == 2013) 
calls_2013[order(calls_2013$month, calls_2013$day, calls_2013$time),]
saveRDS(calls_2013, "~/Documents/GitHub/homelessness-in-the-us/new-york-311/2013-NYC-Service-Requests.Rds")

#2014
calls_2014 <- all_nyc_calls %>%
  filter(year == 2014) 
calls_2014[order(calls_2014$month, calls_2014$day, calls_2014$time),]
saveRDS(calls_2014, "~/Documents/GitHub/homelessness-in-the-us/new-york-311/2014-NYC-Service-Requests.Rds")

#2015
calls_2015 <- all_nyc_calls %>%
  filter(year == 2015) 
calls_2015[order(calls_2015$month, calls_2015$day, calls_2015$time),]
saveRDS(calls_2015, "~/Documents/GitHub/homelessness-in-the-us/new-york-311/2015-NYC-Service-Requests.Rds")

#2016
calls_2016 <- all_nyc_calls %>%
  filter(year == 2016) 
calls_2016[order(calls_2016$month, calls_2016$day, calls_2016$time),]
saveRDS(calls_2016, "~/Documents/GitHub/homelessness-in-the-us/new-york-311/2016-NYC-Service-Requests.Rds")

#2017
calls_2017 <- all_nyc_calls %>%
  filter(year == 2017) 
calls_2017[order(calls_2017$month, calls_2017$day, calls_2017$time),]
saveRDS(calls_2017, "~/Documents/GitHub/homelessness-in-the-us/new-york-311/2017-NYC-Service-Requests.Rds")

#2018
calls_2018 <- all_nyc_calls %>%
  filter(year == 2018) 
calls_2018[order(calls_2018$month, calls_2018$day, calls_2018$time),]
saveRDS(calls_2018, "~/Documents/GitHub/homelessness-in-the-us/new-york-311/2018-NYC-Service-Requests.Rds")

#2019
calls_2019 <- all_nyc_calls %>%
  filter(year == 2019) 
calls_2019[order(calls_2019$month, calls_2019$day, calls_2019$time),]
saveRDS(calls_2019, "~/Documents/GitHub/homelessness-in-the-us/new-york-311/2019-NYC-Service-Requests.Rds")

```

### Load all NYC call sheet
```{r}

#all_nyc_calls <- readRDS("~/Documents/GitHub/homelessness-in-the-us/new-york-311/ALL_NYC-311-Service-Requests.Rds")

```

## Observation values pertaining to homelessness:

Agency:
* DHS

Agency Name:
* Department of Homeless Services
  * In 2010, listed as DHS Advantage Programs, which is a one to two-year rental assistance program for homeless families. In existence 2007-2011. Had work requirements.

Complaint Type:
* Homeless encampment
* Panhandling
* Homeless Street Condition
* Homeless Person Assistance

Location:
* Homeless shelter

Descriptor:
* Homeless Issue
  * Under complaint type "Ferry Complaint"
* 2010 - unenroll from Advantage Program
* Payment Not Received from DHS

## Analyze each year {.tabset}

### 2010

```{r}

#Load data
calls_2010 <- readRDS("~/Documents/GitHub/homelessness-in-the-us/new-york-311/2010-NYC-Service-Requests.Rds")

#calls_2010 %>%
  #parse_date_time(time, '%I:%M %p')
  #strptime(time, "%I:%M %p")
  #format(strptime(time, format = '%I:%M%p)

agency_calls_2010 <- calls_2010 %>%
  group_by(agency) %>%
  summarize(count = n())

complaints_calls_2010 <- calls_2010 %>%
  group_by(complaint_type) %>%
  summarize(count = n())

descriptor_calls_2010 <- calls_2010 %>%
  group_by(descriptor) %>%
  summarize(count = n())

location_calls_2010 <- calls_2010 %>%
  group_by(location_type) %>%
  summarize(count = n())

resolution_calls_2010 <- calls_2010 %>%
  group_by(resolution_description) %>%
  summarize(count = n())

# Relevant types: Homeless Encampment, Panhandling
  
#Take the year-by-year dataframes and only return those with mention of homeless
homeless_2010 <- calls_2010 %>%
  filter_all(any_vars(. %in% c('homeless')))

saveRDS(calls_2010, "~/Documents/GitHub/homelessness-in-the-us/new-york-311/2010-HOMELESS-NYC-Service-Requests.Rds")

```

### 2011

```{r}
#Load data
calls_2011 <- readRDS("~/Documents/GitHub/homelessness-in-the-us/new-york-311/2011-NYC-Service-Requests.Rds")

#Take the year-by-year dataframes and only return those with mention of homeless
homeless_2011 <- calls_2011 %>%
  filter_all(any_vars(. %in% c('homeless')))

saveRDS(calls_2011, "~/Documents/GitHub/homelessness-in-the-us/new-york-311/2011-HOMELESS-NYC-Service-Requests.Rds")

```

### 2012

```{r}
#Load data
calls_2012 <- readRDS("~/Documents/GitHub/homelessness-in-the-us/new-york-311/2012-NYC-Service-Requests.Rds")

#Take the year-by-year dataframes and only return those with mention of homeless
homeless_2012 <- calls_2012 %>%
  filter_all(any_vars(. %in% c('homeless')))

saveRDS(calls_2012, "~/Documents/GitHub/homelessness-in-the-us/new-york-311/2012-HOMELESS-NYC-Service-Requests.Rds")

```

### 2013

```{r}
#Load data
calls_2013 <- readRDS("~/Documents/GitHub/homelessness-in-the-us/new-york-311/2013-NYC-Service-Requests.Rds")

#Take the year-by-year dataframes and only return those with mention of homeless
homeless_2013 <- calls_2013 %>%
  filter_all(any_vars(. %in% c('homeless')))

saveRDS(calls_2013, "~/Documents/GitHub/homelessness-in-the-us/new-york-311/2013-HOMELESS-NYC-Service-Requests.Rds")

```

### 2014

```{r}
#Load data
calls_2014 <- readRDS("~/Documents/GitHub/homelessness-in-the-us/new-york-311/2014-NYC-Service-Requests.Rds")
  
#Take the year-by-year dataframes and only return those with mention of homeless
homeless_2014 <- calls_2014 %>%
  filter_all(any_vars(. %in% c('homeless')))

saveRDS(calls_2014, "~/Documents/GitHub/homelessness-in-the-us/new-york-311/2014-HOMELESS-NYC-Service-Requests.Rds")

```

### 2015

```{r}
#Load data
calls_2015 <- readRDS("~/Documents/GitHub/homelessness-in-the-us/new-york-311/2015-NYC-Service-Requests.Rds")

#Take the year-by-year dataframes and only return those with mention of homeless
homeless_2015 <- calls_2015 %>%
  filter_all(any_vars(. %in% c('homeless')))

saveRDS(calls_2015, "~/Documents/GitHub/homelessness-in-the-us/new-york-311/2015-HOMELESS-NYC-Service-Requests.Rds")

```

### 2016

```{r}
#Load data
calls_2016 <- readRDS("~/Documents/GitHub/homelessness-in-the-us/new-york-311/2016-NYC-Service-Requests.Rds")

#Take the year-by-year dataframes and only return those with mention of homeless
homeless_2016 <- calls_2016 %>%
  filter_all(any_vars(. %in% c('homeless')))

saveRDS(calls_2016, "~/Documents/GitHub/homelessness-in-the-us/new-york-311/2016-HOMELESS-NYC-Service-Requests.Rds")

```

### 2017

```{r}
#Load data
calls_2017 <- readRDS("~/Documents/GitHub/homelessness-in-the-us/new-york-311/2017-NYC-Service-Requests.Rds")

#Take the year-by-year dataframes and only return those with mention of homeless
homeless_2017 <- calls_2017 %>%
  filter_all(any_vars(. %in% c('homeless')))

saveRDS(calls_2017, "~/Documents/GitHub/homelessness-in-the-us/new-york-311/2017-HOMELESS-NYC-Service-Requests.Rds")

```

### 2018

```{r}
#Load data
calls_2018 <- readRDS("~/Documents/GitHub/homelessness-in-the-us/new-york-311/2018-NYC-Service-Requests.Rds")

#Take the year-by-year dataframes and only return those with mention of homeless
homeless_2018 <- calls_2018 %>%
  filter_all(any_vars(. %in% c('homeless')))

saveRDS(calls_2018, "~/Documents/GitHub/homelessness-in-the-us/new-york-311/2018-HOMELESS-NYC-Service-Requests.Rds")

```

### 2019

```{r}
#Load data
calls_2019 <- readRDS("~/Documents/GitHub/homelessness-in-the-us/new-york-311/2019-NYC-Service-Requests.Rds")

#Take the year-by-year dataframes and only return those with mention of homeless
homeless_2019 <- calls_2019 %>%
  filter_all(any_vars(. %in% c('homeless')))

saveRDS(calls_2019, "~/Documents/GitHub/homelessness-in-the-us/new-york-311/2019-HOMELESS-NYC-Service-Requests.Rds")

```
