---
title: "CoC Criminal Survey Analysis"
author: "Theresa Diffendal"
date: "4/23/2020"
output:
  html_document:
    df_print:  paged
    toc: yes
    toc_depth:  3
    toc_float:  yes
    code_folding:  hide
  pdf_document:
    toc:  yes
    toc_depth:  3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This rMarkdown file analyzes criminal charges related to homelessness in 15 states and the District of Columbia as collected by students in Deborah Nelson's JOUR698I class at the University of Maryland's Philip Merrill College of Journalism's Howard Center for Investigative Journalism. Cities in each of the states were examined and codified by whether the law applied to the whole city (written as "citywide"), only applied to particular places ("particular_places"), there was no law ("none"), or the law was unclear ("unclear"). Laws were not examined for every city in each state; cities were examined based on the focus of the class' investigation.

The acronym CoC will be used to mean Continuum of Care, which are citywide or regional areas across which housing and services for the homeless are coordinated. Read more about CoCs [here](https://www.hudexchange.info/programs/coc/).

```{r, warning=FALSE, message=FALSE, echo=FALSE}
## Read in data and clean

##Libraries

#update.packages()

library(rmarkdown)
library(knitr)
library(tidyverse)
library(janitor)
library(lubridate)
library(ggplot2)
library(dplyr)
library(stringr)
library(kableExtra)

## Set working directory
setwd("~/GitHub/homelessness-in-the-us/coc-crim-survey")

crim_survey_initial <- read_csv("coc-crim-survey.csv") %>%
  ##only select rows with data

  ##rename the columns
  rename("name" = "coc-crim-",
         "coc" = X2,
         "city" = X3,
         "public_sleeping_law"  = X10,
         "public_sleeping_link" = X11,
         "public_camping_law" = X13,
         "public_camping_link" = X14,
         "public_sitting_law" = X16,
         "public_sitting_link" = X17,
         "vehicle_living_law" = X19,
         "vehicle_living_link" = X20,
         "loitering_law" = X22,
         "loitering_link" = X23,
         "begging_law" = X25,
         "begging_link" = X26,
         "food_sharing_law" = X28,
         "food_sharing_link" = X29,
         "public_urination_law" = X31,
         "public_urination_link" = X32,
         "public_intoxication_law" = X34,
         "public_intoxication_link" = X35,
         "misc_law" = X37,
         "misc_link" = X38) %>%

  ##clean column names
  clean_names() %>%
  ##change all observations to lower case
  mutate_all(funs(tolower)) %>%
  mutate(coc = str_sub(coc, 4,-1)) %>%
  rename("living_in_vehicle" = "lodging_living_sleeping_in_vehicles_or_parking_a_vehicle_used_as_a_lodging")

crim_survey_initial[crim_survey_initial == "county"] <- "citywide"

crim_survey <- crim_survey_initial %>%
  ##create columns to summarize how many laws are citywide, particular, none or unclear in each row
  mutate(citywide = rowSums(crim_survey_initial == "citywide", na.rm=TRUE),   
         particular_places = rowSums(crim_survey_initial == "only in particular places", na.rm=TRUE),
         none = rowSums(crim_survey_initial == "none", na.rm=TRUE),
         unclear = rowSums(crim_survey_initial == "unclear", na.rm=TRUE)) %>%

  ##get rid of columns we don't need to analyze
  filter(complete == "yes",
         coc != "state",
         c3 == "yes",
         meet_definition == "yes") %>%
  select(-name, -complete, -new_reporting_or_nlchp_fact_check, -c3, -meet_definition, -contains ("link"), -contains("law")) %>%
  ##reorder columns with mutated columns first
  select(3,1,2,13:16,4:12)
```

## Overall Law Information

### Complete dataset

Below is the complete dataset which journalists compiled and is the source for this data analysis.
```{r}
crim_survey <- crim_survey %>%
  mutate(total_laws = rowSums(crim_survey[,c("citywide", "particular_places")], na.rm=TRUE)) %>%
  select(1:3,17,4:16) %>%
  arrange(state)


kable(crim_survey) %>%
  column_spec(2:ncol(crim_survey), border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "350px")
```


Based on pre-reporting, the following laws were examined as they disproportionately impact homeless individuals. Homeless individuals often end up with multiple criminal charges and exorbitant through the criminalization of necessary acts such as begging, sharing food or camping in public. The list below contains the nine laws.
```{r}
colnames(crim_survey[9:17])
```

There are more citywide laws against public intoxication and public urination than any other crimes, with 34 each. Begging is the third act most criminalized citywide with laws in 30 cities.
```{r}
law_type_totals <- crim_survey %>%
  select(9:17) %>%
  gather() %>%
  group_by(key, value) %>%
  filter(value != "unclear",
         value != "none") %>%
  count()

kable(law_type_totals) %>%
  column_spec(2:ncol(law_type_totals), border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "350px")
```

Public intoxication just tops camping in public for the dubious record of most criminalized. Food sharing bans are the least popular, only existing in four cities.
```{r}
law_totals <- law_type_totals %>%
  group_by(key) %>%
  summarise(count=sum(n)) %>%
  arrange(desc(count))
print(law_totals)
```

Most of the laws examined are applicable across the city, which is only slightly greater than the amount of laws that are not on the books. Together, "citywide" and "none" account for over 70% of the laws examined. 21% of laws are only applicable in particular places while 9% of the laws were unclear.

While laws applicable in particular places represent a fifth of the laws examined, the majority of the time it appears that laws either apply citywide or do not exist.
```{r}
overall_laws <- crim_survey %>%
  select(citywide, particular_places, unclear, none) %>%
  colSums() %>%
  as.data.frame(, keep.rownames=TRUE) %>%
  tibble::rownames_to_column() %>%
  rename("count" = ".") %>%
  mutate(pct = (count / sum(count))*100)

print(overall_laws)
```

## State and City Information

* 15 states are examined: Alaska, California, Florida, Georgia, Hawaii, Kentucky, Louisiana, Massachusetts, Maryland, Missouri, Nevada, New York, North Carolina, Oregon and Washington; as well as the District of Columbia.

Abbreviations of every state:
```{r}
#Get the abbreviations of all examined states
states <- as.vector(crim_survey$state)
print(unique(states))
```
Number of states + D.C.:
```{r}
# See how many states there are; answer: 15 + D.C.
length(unique(states))
```

Number of cities analyzed in each state
```{r}
#How many cities were analyzed in each state
crim_survey_states <- crim_survey %>%
  group_by(state) %>%
  summarise(city_count=n())
print(crim_survey_states)
```

Total cities examined:

```{r}
print(sum(crim_survey_states$city_count))
```

Names of cities examined in each state:
```{r}
# See each city examined in each state
crim_survey_cities <- crim_survey %>%
  group_by(state) %>%
  select(state, city) %>%
  arrange(state)

## all the kable code creates an interactive, scrollable table in the .html output
kable(crim_survey_cities) %>%
  column_spec(2, border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "250px")
```

## City Analysis

* Oakland has the most with 9 laws. This means that Oakland has every possible examined law on the books.
* Oakland, California has the greatest number of laws that disproportionately affect homeless individuals with 9, which means they have a law banning each examined act. San Francisco and Los Angeles and Atlanta, Georgia are tied for a close seconds with 8 laws.
  + California cities account for three of the top four cities with the most laws.
* Cities in California and Florida also have the second fewest laws that mostly impact homeless individuals: Ione California and Lake City, Florida only have 2 laws.
* The city of Waldorf, Md. has the fewest laws with only 1. 
  + Waldorf's one law outlaws begging citywide.
```{r}
## create new data set
city_totals <- crim_survey %>%

  ## group by state so totals are by state
  group_by(state, city) %>%

  ## sum all indications that a law exists (so not including none or unclear)
  summarise(total_laws = sum(citywide + particular_places),
            "citywide"=citywide,
            "particular_places"=particular_places,
            "unclear"=unclear,
            "none"=none) %>%

  ## put the state with the most laws first
  arrange(desc(total_laws))

kable(city_totals) %>%
  column_spec (2:ncol(city_totals),border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "250px")
```

### Type of laws by City
```{r}
type_by_city <- crim_survey %>%
  select(3,9:17) %>%
  group_by(city) %>%
  naniar::replace_with_na_all(condition = ~.x == "unclear") %>%
  naniar::replace_with_na_all(condition = ~.x == "none")

type_by_city[type_by_city == "citywide"] <- 1
type_by_city[type_by_city == "only in particular places"] <- 1
type_by_city[type_by_city == "county"] <- 1

type_by_city[,2:10] <- sapply(type_by_city[,2:10], as.numeric)

type_by_city <- aggregate(type_by_city[,2:10], by=list(type_by_city$city), FUN=sum, na.rm=TRUE) %>%
  rename("city" = "Group.1")

kable(type_by_city) %>%
  column_spec(2:ncol(type_by_city),border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "300px")
```

## Types of Laws by State

### Which state has the most laws?

* California comes in first with 109 laws and is the only examined state with over 50 laws.
* Kentucky has the fewest laws with only 4
* Maryland is the only state in which more than one city was examined but, due to the total number of laws, fell behind multiple states with only one city.
* The top six states - California, Florida, Massachusetts, Oregon, Georgia and Hawaii - were the only states (besides Maryland) in which more than one CoC was examined.
```{r}
## create new data set
state_totals <- crim_survey %>%

  ## group by state so totals are by state
  group_by(state) %>%

  ## sum all indications that a law exists (so not including none or unclear)
  summarise(total_laws = sum(citywide + particular_places)) %>%

  ## get rid of other columns
  select(state, total_laws) %>%

  ## put the state with the most laws first
  arrange(desc(total_laws))

kable(state_totals) %>%
  column_spec (2:ncol(state_totals),border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "250px")
```

### What acts do each state outlaw?

```{r}
type_by_state <- crim_survey %>%
  select(1,9:17) %>%
  group_by(state) %>%
  naniar::replace_with_na_all(condition = ~.x == "unclear") %>%
  naniar::replace_with_na_all(condition = ~.x == "none")

type_by_state[type_by_state == "citywide"] <- 1
type_by_state[type_by_state == "only in particular places"] <- 1
type_by_state[type_by_state == "county"] <- 1

type_by_state[,2:10] <- sapply(type_by_state[,2:10], as.numeric)

type_by_state <- aggregate(type_by_state[,2:10], by=list(type_by_state$state), FUN=sum, na.rm=TRUE) %>%
  rename("state" = "Group.1")

kable(type_by_state) %>%
  column_spec (2:ncol(type_by_state),border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "300px")
```


**Notes:**

* There are more citywide laws than any other kind, followed by no law, particular places  and unclear.
* In 12 of the 15 states (and DC), fifty percent or less of the laws analyzed are citywide.
* Hawaii and Ga. had the lowest percentage of citywide laws with only 16.7%
  + Hawaii also had the greatest share of "unclear" laws, accounting for seven (27.8%) of its examined laws
* Over half  of the examined laws in New York City (55.56%) and Georgia (61.1%) are only applicable in particular places. Over in New Orleans, 44.4% of the laws cover a particular place.
* Missouri and Alaska have no laws only applicable in particular places.

```{r}
all_state <- crim_survey %>%

  #only select the columns with sums and calculations
  select(c(1,3,5:8)) %>%

  #group by state in this dataset instead of city in the previous sets
  group_by(state) %>%

  #sum the number of citywide, particular, none, and unclear laws for each state
  summarise(sum_state_citywide = sum(citywide),
            sum_state_particular_places = sum(particular_places),
            sum_state_none = sum(none),
            sum_state_unclear = sum(unclear))

all_state <- all_state %>%

  #calculate what percentage of the state's laws are citywide, particular, none, unclear
  mutate(percent_state_citywide = (sum_state_citywide / rowSums(all_state[,c(2:5)]) * 100),
         percent_state_particular_places = (sum_state_particular_places / rowSums(all_state[,c(2:5)]) * 100),
         percent_state_none = (sum_state_none / rowSums(all_state[,c(2:5)]) * 100),
         percent_state_unclear = (sum_state_unclear / rowSums(all_state[,c(2:5)]) * 100)) %>%

  #add a row with totals
  janitor::adorn_totals("row")

kable(all_state) %>%
  column_spec (2:ncol(all_state),border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "300px")

```

## State by State Data {.tabset}

### Alaska

Overview:

* While Alaska will allow people to use their vehicles as lodging, there are citywide laws against sitting or laying in public, sleeping in public, and camping in public.
* There is no food sharing ban but there is a citywide law against begging
```{r}
alaska <- crim_survey %>%
  subset(state %in% "ak")

kable(alaska) %>%
  column_spec(3:ncol(alaska), border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "200px")
```

### California

```{r}
california <- type_by_state %>%
  filter(state == "ca")

california <- california %>%
  mutate(total_laws = sum(california[,c(2:10)], na.rm=TRUE)) %>%
  select(1,11,2:10)

kable(california) %>%
  column_spec (2:ncol(california),border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "100px")
```

```{r}
california_cities <- crim_survey %>%
  subset(state %in% "ca")

kable(california_cities) %>%
  column_spec (3:ncol(california_cities),border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "500px")
```

### DC

```{r}
dc <- crim_survey %>%
  subset(state %in% "dc")

kable(dc) %>%
  column_spec (3:ncol(dc),border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "250px")
```

### Florida
* Journalists examined the lawsin 9 cities in Florida, including Daytona Beach, Gainesville, Fort Walton Beach and St. Petersburg.
* Florida cities ban camping in public more than any other act, with laws on the books in 8 cities. Living in a vehicle and loitering or vagrancy round out the top three most criminalized with laws in 7 cities.
* The city of Daytona Beach criminalizes seven of the nine acts identified as disproportionately targeting homeless individuals and acts of homelessness (it does not ban sitting or laying in public and sharing food). 
  + All the laws are citywide except for one, camping in public. The ban on camping in public is only applicable in particular places, suggesting that some areas in the city are purposefully reserved for public camping, though whether those campers are intended to be tourists or homeless individuals is unknown.
  + Daytona does permit sitting and laying in public, but the lack of allowance for sleeping either in public or a car leaves no exception for those who remain homeless when night falls.
* Lake City in the Columbia CoC is the most lenient, with citywide laws only against public intoxication and vagrancy and loitering. Though an encampment might violate the loitering and vagrancy law, both of these acts are also committed by vast swaths of the population and thus target homeless populations less than a ban on sleeping in public or sharing food.
```{r}
florida <- type_by_state %>%
  filter(state == "fl")

florida <- florida %>%
  mutate(total_laws = sum(florida[,c(2:10)], na.rm=TRUE)) %>%
  select(1,11,2:10)

kable(florida) %>%
  column_spec (2:ncol(florida),border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "100px")
```

```{r}
florida_cities <- crim_survey %>%
  subset(state %in% "fl")

kable(florida_cities) %>%
  column_spec (3:ncol(florida_cities),border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "500px")
```

### Hawaii

```{r}
hawaii <- type_by_state %>%
  filter(state == "hi")

hawaii <- hawaii %>%
  mutate(total_laws = sum(hawaii[,c(2:10)], na.rm=TRUE)) %>%
  select(1,11,2:10)

kable(hawaii) %>%
  column_spec (2:ncol(hawaii),border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "100px")
```

```{r}
hawaii_cities <- crim_survey %>%
  subset(state %in% "hi")

kable(hawaii_cities) %>%
  column_spec (3:ncol(hawaii_cities),border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "250px")
```

### Kentucky

```{r}
kentucky <- crim_survey %>%
  subset(state %in% "ky")

kable(kentucky) %>%
  column_spec (3:ncol(kentucky),border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "250px")
```

### Louisiana

```{r}
louisiana <- crim_survey %>%
 subset(state %in% "la")

kable(louisiana) %>%
 column_spec (3:ncol(louisiana),border_left = T, border_right = F) %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
scroll_box(height = "250px")
```

### Maryland

```{r}
maryland <- type_by_state %>%
  filter(state == "md")

maryland <- maryland %>%
  mutate(total_laws = sum(maryland[,c(2:10)], na.rm=TRUE)) %>%
  select(1,11,2:10)

kable(maryland) %>%
  column_spec (2:ncol(maryland),border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "100px")
```

```{r}
maryland_cities <- crim_survey %>%
  subset(state %in% "md")

kable(maryland_cities) %>%
  column_spec (3:ncol(maryland_cities),border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "250px")
```

### Massachusetts

```{r}
massachusetts <- type_by_state %>%
  filter(state == "ma")

massachusetts <- massachusetts %>%
  mutate(total_laws = sum(massachusetts[,c(2:10)], na.rm=TRUE)) %>%
  select(1,11,2:10)

kable(massachusetts) %>%
  column_spec (2:ncol(massachusetts),border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "100px")
```

```{r}
massachusetts_cities <- crim_survey %>%
  subset(state %in% "ma")

kable(massachusetts_cities) %>%
  column_spec (3:ncol(massachusetts_cities),border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "300px")
```

### New York

```{r}
new_york <- crim_survey %>%
  subset(state %in% "ny")

kable(new_york) %>%
  column_spec (3:ncol(new_york),border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "175px")
```

### North Carolina

```{r}
north_carolina <- crim_survey %>%
  subset(state %in% "nc")

kable(north_carolina) %>%
  column_spec (3:ncol(north_carolina),border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "250px")
```

### Oregon

```{r}
oregon <- type_by_state %>%
  filter(state == "or")

oregon <- oregon %>%
  mutate(total_laws = sum(oregon[,c(2:10)], na.rm=TRUE)) %>%
  select(1,11,2:10)

kable(oregon) %>%
  column_spec (2:ncol(oregon),border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "100px")
```

```{r}
oregon_cities <- crim_survey %>%
  subset(state %in% "or")

kable(oregon_cities) %>%
  column_spec (3:ncol(oregon_cities),border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "250px")
```

## Encampment Analysis

The following is an analysis of laws only related to homeless encampments: sleeping in public, camping in public, sitting/laying in public, and lodging/living/sleeping in vehicles or parking a vehicle used as a lodging.

### Encampment Law Overview

* Individuals who either cannot or do not want to find room at a homeless shelter will often turn to homeless encampments. Though the lodgings are impermanent, lack utilities and hardly protect against the elements, living in a community engenders feelings of belonging and may seem safer than surviving independently. But even as encampment populations increase nation-wide due to a lack of shelters and beds, so too do the number of laws against camping in public.
* Homeless encampments are threatened by laws banning camping, sitting, laying or sleeping in public or living in a vehicle. Yet many cities seek to downsize their homeless populations by enforcing such laws, with the consequence of sending many homeless people into the criminal justice system for commiting necessary acts to survive while homeless.
* In an analysis of nine laws in 53 cities across the country that disproportionately impact homeless individuals, only public intoxication was criminalized more than camping in public. A little less than half of cities only banned public camping in particular places, but it’s unknown if there are allowances for homeless encampments.
* More than half of all examined cities had bans on some aspect of homeless encampments defined as camping in public, sitting or laying in public, sleeping in public or living in a vehicle. 
* Eleven cities had laws against every aspect, including Marysville and Oakland, California, where all four laws are applicable citywide.
* Six cities, however, such as Baltimore City, St. Louis and Lake City, Florida, do not criminalize any basic aspects of residing in a homeless encampment.
* Most states had the most bans on either camping or sitting and laying in public, except California, which seems to be primarily concerned with people living out of vehicles.

```{r}
encampment_totals <- law_totals %>%
  filter(key == "camping_in_public" |
           key == "sitting_laying_in_public" |
           key == "sleeping_in_public" |
           key == "living_in_vehicle")

kable(encampment_totals) %>%
  column_spec (2:ncol(encampment_totals),border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "250px")

encampment_type_totals <- law_type_totals %>%
  filter(key == "camping_in_public" |
           key == "sitting_laying_in_public" |
           key == "sleeping_in_public" |
           key == "living_in_vehicle")

kable(encampment_type_totals) %>%
  column_spec (2:ncol(encampment_type_totals),border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "250px")
```

### Encampment Laws by State and City

```{r}
encampments <- crim_survey_initial %>%
  ##get rid of columns we don't need to analyze
   filter(complete == "yes",
         coc != "state",
         c3 == "yes",
         meet_definition == "yes") %>%
  select(-name, -complete, -new_reporting_or_nlchp_fact_check, -c3, -meet_definition, -contains ("link"), -contains("law")) %>%
  select(3,1,2,4:7)

encampments <- encampments %>%
  mutate(citywide = rowSums(encampments == "citywide", na.rm=TRUE),   
         particular_places = rowSums(encampments == "only in particular places", na.rm=TRUE),
         none = rowSums(encampments == "none", na.rm=TRUE),
         unclear = rowSums(encampments == "unclear", na.rm=TRUE))

encampments <- encampments %>%  
  mutate(total_laws = rowSums(encampments[,c("citywide", "particular_places")], na.rm=TRUE)) %>%
  select(1:3,12,8:11,4:7) %>%
  arrange(state)

kable(encampments) %>%
  column_spec (3:ncol(encampments),border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "400px")
```

### Encampment Laws by State
```{r}
encampment_state <- encampments %>%
  select(1,4:6,9:12) %>%
  group_by(state) %>%
  naniar::replace_with_na_all(condition = ~.x == "unclear") %>%
  naniar::replace_with_na_all(condition = ~.x == "none")

encampment_state[encampment_state == "citywide"] <- 1
encampment_state[encampment_state == "only in particular places"] <- 1
encampment_state[encampment_state == "county"] <- 1

encampment_state[,2:8] <- sapply(encampment_state[,2:8], as.numeric)
  encampment_state <- aggregate(encampment_state[,2:8], by=list(encampment_state$state), FUN=sum, na.rm=TRUE) %>%
  rename("state" = "Group.1")

kable(encampment_state) %>%
  column_spec (3:ncol(encampment_state),border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "400px")
```

## Cross Examine CoCs with Homeless Counts

Read in HUD PIT counts and join to crim survey
```{r}
pit <- read_csv("~/GitHub/homelessness-in-the-us/data/output-data/pit-hic-zillow.csv") %>%
  rename(coc = coc_name)

homeless_crim <- left_join(crim_survey, pit, by="coc")
```

```{r, include=FALSE}
## Write csv files for adam with cocs as rows and Y/N laws as columns

#type_by_coc_adam <- type_by_coc
#type_by_coc_adam[type_by_coc_adam == "1"] <- "Y"
#type_by_coc_adam[type_by_coc_adam == "0"] <- "N"
#view(type_by_coc_adam)

#write.csv(type_by_coc_adam, "crim-survey-adam.csv")
```

```{r, include=FALSE}
## Write csv files for all state summaries
#setwd("~/Documents/GitHub/homelessness-in-the-us/coc-crim-survey")
#write.csv(alaska, "alaska-summary.csv")
#write.csv(california, "california-summary.csv")
#write.csv(dc, "dc-summary.csv")
#write.csv(florida, "florida-summary.csv")
#write.csv(hawaii, "hawaii-summary.csv")
#write.csv(kentucky, "kentucky-summary.csv")
#write.csv(louisiana, "louisiana-summary.csv")
#write.csv(maryland, "maryland-summary.csv")
#write.csv(massachusetts, "massachusetts-summary.csv")
#write.csv(new_york, "new-york-summary.csv")
#write.csv(north_carolina, "north-carolina-summary.csv")
#write.csv(oregon, "oregon-summary.csv")
```

