---
title: "CoC Criminal Survey Analysis"
author: "Theresa Diffendal"
date: "4/13/2020"
output:
  html_document:
    df_print:  paged
    toc: yes
    toc_depth:  3
    toc_float:  yes
    code_folding:  hide
  pdf_document:
    toc:  yes
    toc_depth:  3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This rMarkdown file analyzes criminal charges related to homelessness in 15 states and the District of Columbia as collected by students in Deborah Nelson's JOUR698I class at the University of Maryland's Philip Merrill College of Journalism's Howard Center for Investigative Journalism. Cities in each of the states were examined and codified by whether the law applied to the whole city (written as "citywide"), only applied to particular places ("particular_places"), there was no law ("none"), or the law was unclear ("unclear"). Laws were not examined for every city in each state; cities were examined based on the focus of the class' investigation. 

The acronym CoC will be used to mean Continuum of Care, which are citywide or regional areas across which housing and services for the homeless are coordinated. Read more about CoCs [here](https://www.hudexchange.info/programs/coc/).

```{r, warning=FALSE, message=FALSE, echo=FALSE}
## Read in data and clean

##Libraries

#update.packages()

library(rmarkdown)
library(knitr)
library(tidyverse)
library(janitor)
library(lubridate)
library(ggplot2)
library(dplyr)
library(stringr)
library(kableExtra)

## Set working directory
setwd("~/GitHub/homelessness-in-the-us/coc-crim-survey")

crim_survey_initial <- read_csv("coc-crim-survey.csv") %>%
  ##only select rows with data
  
  ##rename the columns
  rename("name" = "coc-crim-",
         "coc" = X2,
         "city" = X3,
         "public_sleeping_law"  = X10,
         "public_sleeping_link" = X11,
         "public_camping_law" = X13,
         "public_camping_link" = X14,
         "public_sitting_law" = X16,
         "public_sitting_link" = X17,
         "vehicle_living_law" = X19,
         "vehicle_living_link" = X20,
         "loitering_law" = X22,
         "loitering_link" = X23,
         "begging_law" = X25,
         "begging_link" = X26,
         "food_sharing_law" = X28,
         "food_sharing_link" = X29,
         "public_urination_law" = X31,
         "public_urination_link" = X32,
         "public_intoxication_law" = X34,
         "public_intoxication_link" = X35,
         "misc_law" = X37,
         "misc_link" = X38) %>%
  
  ##clean column names
  clean_names() %>%
  ##change all observations to lower case
  mutate_all(funs(tolower)) %>%
  mutate(coc = str_sub(coc, 4,-1)) %>%
  rename("living_in_vehicle" = "lodging_living_sleeping_in_vehicles_or_parking_a_vehicle_used_as_a_lodging")

crim_survey_initial[crim_survey_initial == "county"] <- "citywide"

crim_survey <- crim_survey_initial %>%
  ##create columns to summarize how many laws are citywide, particular, none or unclear in each row
  mutate(citywide = rowSums(crim_survey_initial == "citywide", na.rm=TRUE),   
         particular_places = rowSums(crim_survey_initial == "only in particular places", na.rm=TRUE),
         none = rowSums(crim_survey_initial == "none", na.rm=TRUE),
         unclear = rowSums(crim_survey_initial == "unclear", na.rm=TRUE)) %>%
  
  ##get rid of columns we don't need to analyze
  filter(complete == "yes",
         coc != "state",
         c3 == "yes",
         meet_definition == "yes") %>%
  select(-name, -complete, -new_reporting_or_nlchp_fact_check, -c3, -meet_definition, -contains ("link"), -contains("law")) %>%
  ##reorder columns with mutated columns first
  select(3,1,2,13:16,4:12)
```
  
## Overall Law Information

### Complete dataset
  
Below is the complete dataset which journalists compiled and is the source for this data analysis.
```{r}
crim_survey <- crim_survey %>%
  mutate(total_laws = rowSums(crim_survey[,c("citywide", "particular_places")], na.rm=TRUE)) %>%
  select(1:3,17,4:16) %>%
  arrange(state)
  

kable(crim_survey) %>%
  column_spec(2:ncol(crim_survey), border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "350px")
```
  
    
Based on pre-reporting, the following laws were examined as they disproportionately impact homeless individuals. Homeless individuals often end up with multiple criminal charges and exorbitant through the criminalization of necessary acts such as begging, sharing food or camping in public. The list below contains the nine laws 
```{r}
colnames(crim_survey[9:17])
```

34 of the 41 laws against public intoxication apply citywide, the most citywide-applicable laws for any crime. Public urination is criminalized citywide in 33 cities, while begging is the third act most criminalized citywide with laws in 29 cities.
```{r}
law_type_totals <- crim_survey %>%
  select(9:17) %>%
  gather() %>%
  group_by(key, value) %>%
  filter(value != "unclear",
         value != "none") %>%
  count() 

kable(law_type_totals) %>%
  column_spec(2:ncol(law_type_totals), border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "350px")
```
  
Public intoxication just tops camping in public for the dubious record of most criminalized. Food sharing bans are the least popular, only existing in four cities. 
```{r}
law_totals <- law_type_totals %>%
  group_by(key) %>%
  summarise(count=sum(n)) %>%
  arrange(desc(count))
print(law_totals)
```
  
Most of the laws examined are applicable across the city, which is only slightly greater than the amount of laws that are not on the books. Together, "citywide" and "none" account for over 70% of the laws examined. 20% of laws are only applicable in particular places while 8% of the laws were unclear.

While laws applicable in particular places represent a fifth of the laws examined, the majority of the time it appears that laws either apply citywide or do not exist.
```{r}
overall_laws <- crim_survey %>%
  select(citywide, particular_places, unclear, none) %>%
  colSums() %>%
  as.data.frame(, keep.rownames=TRUE) %>%
  tibble::rownames_to_column() %>%
  rename("count" = ".") %>%
  mutate(pct = (count / sum(count))*100) 

print(overall_laws)
```

## State and City Information

* 15 states are examined: Alaska, California, Florida, Georgia, Hawaii, Kentucky, Louisiana, Massachusetts, Maryland, Missouri, Nevada, New York, North Carolina, Oregon and Washington; as well as the District of Columbia.
  
Abbreviations of every state:
```{r}
#Get the abbreviations of all examined states
states <- as.vector(crim_survey$state)
print(unique(states))
```
Number of states + D.C.:
```{r}
# See how many states there are; answer: 13 + D.C.
length(unique(states))
```

Number of cities analyzed in each state
```{r}
#How many cities were analyzed in each state
crim_survey_states <- crim_survey %>%
  group_by(state) %>%
  summarise(city_count=n())
print(crim_survey_states)
```
  
Names of cities examined in each state
```{r}
# See each city examined in each state
crim_survey_cities <- crim_survey %>%
  group_by(state) %>%
  select(state, city) %>%
  arrange(state)

## all the kable code creates an interactive, scrollable table in the .html output
kable(crim_survey_cities) %>%
  column_spec(2, border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "250px")
```

## Continuum of Care Analysis

### CoCs by State

A list of CoCs in each state
```{r}
#list cocs in each state
cocs <- crim_survey %>%
  select(state, coc) %>%
  distinct(state, coc) %>%
  group_by(state) %>%
  arrange(state, coc)

kable(cocs) %>%
  column_spec(2, border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "300px")
```
  
The number of CoCs in each state
```{r}
#number cocs in each state
cocs <- cocs %>%
  group_by(state) %>%
  summarise(count=n())

kable(cocs) %>%
  column_spec(2, border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "250px")
```

### **Location of laws in CoCs:**

* California CoC Oakland/Alameda has the greatest number of laws that disproportionately affect homeless individuals with 9, which means they have a law banning each examined act. California's San Francisco CoC and Atlanta, Georgia's CoC are close seconds with 8 laws.
* CoCs in California and Florida also have the second fewest laws that mostly impact homeless individuals. Amador/Calaveras/Tuolumne/Mariposa and Humboldt County CoCs in California and Columbia/Hamilton/Lafayette/Suwanee Coc in Florida only have 2 laws.
* A Maryland CoC consisting of Charles, Calvert and St. Mary's counties nq  ` x has only 1 law. It outlaws begging citywide.
```{r}
## create new data set
coc_totals <- crim_survey %>%
  select(coc, citywide, particular_places, unclear, none) %>%
  ## group by coc so totals are by coc
  group_by(coc)
  
coc_totals <- aggregate(coc_totals[,2:5], by=list(coc_totals$coc), FUN=sum, na.rm=TRUE) %>%
  rename("coc" = "Group.1") %>%
  group_by(coc) %>%
  mutate(total_laws = sum(citywide + particular_places)) %>%
  select(1, 6, 2:5) %>%
  ## put the state with the most laws first
  arrange(desc(total_laws))

kable(coc_totals) %>%
  column_spec (1:ncol(coc_totals),border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "300px")
```
  
### Type of laws by CoC
```{r}
type_by_coc <- crim_survey %>%
  select(2,9:17) %>%
  group_by(coc) %>%
  naniar::replace_with_na_all(condition = ~.x == "unclear") %>%
  naniar::replace_with_na_all(condition = ~.x == "none")

type_by_coc[type_by_coc == "citywide"] <- 1
type_by_coc[type_by_coc == "only in particular places"] <- 1
type_by_coc[type_by_coc == "county"] <- 1

type_by_coc[,2:10] <- sapply(type_by_coc[,2:10], as.numeric)

type_by_coc <- aggregate(type_by_coc[,2:10], by=list(type_by_coc$coc), FUN=sum, na.rm=TRUE) %>%
  rename("coc" = "Group.1")

kable(type_by_coc) %>%
  column_spec(2:ncol(type_by_coc),border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "300px")
```
  
## City Analysis

* The cities of Waldorf, Md. and Worcester, Mass. have the fewest laws with only 1. Lake City, Ione and Eureka round out the bottom five with two laws each.
* Oakland has the most with 9 laws. This means that Oakland has every possible examined law on the books.
* No state has eight laws, but 10 states have seven laws.
```{r}
## create new data set
city_totals <- crim_survey %>%
  
  ## group by state so totals are by state
  group_by(state, city) %>%
  
  ## sum all indications that a law exists (so not including none or unclear)
  summarise(total_laws = sum(citywide + particular_places),
            "citywide"=citywide, 
            "particular_places"=particular_places, 
            "unclear"=unclear, 
            "none"=none) %>%
  
  ## put the state with the most laws first
  arrange(desc(total_laws))

kable(city_totals) %>%
  column_spec (2:ncol(city_totals),border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "250px")
```
  
## Which state has the most laws?

* California comes in first with 115 laws and is the only examined state with over 100 laws.
* Kentucky has the fewest laws with only 4
* The top six states - California, Florida, Oregon, Massachusetts, Hawaii and Maryland - were the only states in which more than one CoC was examined.
```{r}
## create new data set
state_totals <- crim_survey %>%
  
  ## group by state so totals are by state
  group_by(state) %>%
  
  ## sum all indications that a law exists (so not including none or unclear)
  summarise(total_laws = sum(citywide + particular_places)) %>%
  
  ## get rid of other columns
  select(state, total_laws) %>%
  
  ## put the state with the most laws first
  arrange(desc(total_laws))

kable(state_totals) %>%
  column_spec (2:ncol(state_totals),border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "250px")
```

## Types of Laws by State

### What acts do each state outlaw?

```{r}
type_by_state <- crim_survey %>%
  select(1,9:17) %>%
  group_by(state) %>%
  naniar::replace_with_na_all(condition = ~.x == "unclear") %>%
  naniar::replace_with_na_all(condition = ~.x == "none")

type_by_state[type_by_state == "citywide"] <- 1
type_by_state[type_by_state == "only in particular places"] <- 1
type_by_state[type_by_state == "county"] <- 1

type_by_state[,2:10] <- sapply(type_by_state[,2:10], as.numeric)

type_by_state <- aggregate(type_by_state[,2:10], by=list(type_by_state$state), FUN=sum, na.rm=TRUE) %>%
  rename("state" = "Group.1")

kable(type_by_state) %>%
  column_spec (2:ncol(type_by_state),border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "300px")
```


**Notes:**

* There are more citywide laws than any other kind , followed by no law, particular places  and unclear.
* In 10 of the 14 states (and DC), fifty percent or less of the laws analyzed are citywide.
* Hawaii had the lowest percentage citywide laws, accounting for only 3 (16.7%) of the 18 HI laws recorded
  + Hawaii also had the most "unclear" laws, accounting for seven (27.8%) of its examined laws
* Over half (55.56%) of the examined laws in New York City are only applicable in particular places. Over in New Orleans, 44.4% of the laws cover a particular place.
* Massachusetts has no laws only applicable in particular places. 

```{r}
all_state <- crim_survey %>%

  #only select the columns with sums and calculations
  select(c(1,3,5:8)) %>%
  
  #group by state in this dataset instead of city in the previous sets
  group_by(state) %>%
  
  #sum the number of citywide, particular, none, and unclear laws for each state
  summarise(sum_state_citywide = sum(citywide),
            sum_state_particular_places = sum(particular_places),
            sum_state_none = sum(none),
            sum_state_unclear = sum(unclear)) 

all_state <- all_state %>%
  
  #calculate what percentage of the state's laws are citywide, particular, none, unclear
  mutate(percent_state_citywide = (sum_state_citywide / rowSums(all_state[,c(2:5)]) * 100),
         percent_state_particular_places = (sum_state_particular_places / rowSums(all_state[,c(2:5)]) * 100),
         percent_state_none = (sum_state_none / rowSums(all_state[,c(2:5)]) * 100),
         percent_state_unclear = (sum_state_unclear / rowSums(all_state[,c(2:5)]) * 100)) %>%
  
  #add a row with totals
  janitor::adorn_totals("row")

kable(all_state) %>%
  column_spec (2:ncol(all_state),border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "300px")

```

## State by State Data {.tabset}

### Alaska

Overview: 

* Alaska has one law against sleeping in public that is county-wide, the only county law in this criminal survey analysis
+ However, sitting, laying or camping in public is not allowed citywide
* There is no food sharing ban but there is a citywide law against begging
```{r}
alaska <- crim_survey %>%
  subset(state %in% "ak") 

kable(alaska) %>%
  column_spec(3:ncol(alaska), border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "200px")
```

### California

```{r}
california <- type_by_state %>%
  filter(state == "ca") 

california <- california %>%
  mutate(total_laws = sum(california[,c(2:10)], na.rm=TRUE)) %>%
  select(1,11,2:10)

kable(california) %>%
  column_spec (2:ncol(california),border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "100px")
```

```{r}
california_cities <- crim_survey %>%
  subset(state %in% "ca")

kable(california_cities) %>%
  column_spec (3:ncol(california_cities),border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "500px")
```

### DC

```{r}
dc <- crim_survey %>%
  subset(state %in% "dc")

kable(dc) %>%
  column_spec (3:ncol(dc),border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "250px")
```

### Florida

```{r}
florida <- type_by_state %>%
  filter(state == "fl") 

florida <- florida %>%
  mutate(total_laws = sum(florida[,c(2:10)], na.rm=TRUE)) %>%
  select(1,11,2:10)

kable(florida) %>%
  column_spec (2:ncol(florida),border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "100px")
```

```{r}
florida_cities <- crim_survey %>%
  subset(state %in% "fl")

kable(florida_cities) %>%
  column_spec (3:ncol(florida_cities),border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "500px")
```

### Hawaii

```{r}
hawaii <- type_by_state %>%
  filter(state == "hi") 

hawaii <- hawaii %>%
  mutate(total_laws = sum(hawaii[,c(2:10)], na.rm=TRUE)) %>%
  select(1,11,2:10)

kable(hawaii) %>%
  column_spec (2:ncol(hawaii),border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "100px")
```

```{r}
hawaii_cities <- crim_survey %>%
  subset(state %in% "hi")

kable(hawaii_cities) %>%
  column_spec (3:ncol(hawaii_cities),border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "250px")
```

### Kentucky

```{r}
kentucky <- crim_survey %>%
  subset(state %in% "ky")

kable(kentucky) %>%
  column_spec (3:ncol(kentucky),border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "250px")
```

### Louisiana

```{r}
louisiana <- crim_survey %>%
 subset(state %in% "la")

kable(louisiana) %>%
 column_spec (3:ncol(louisiana),border_left = T, border_right = F) %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
scroll_box(height = "250px")
```

### Maryland

```{r}
maryland <- type_by_state %>%
  filter(state == "md") 

maryland <- maryland %>%
  mutate(total_laws = sum(maryland[,c(2:10)], na.rm=TRUE)) %>%
  select(1,11,2:10)

kable(maryland) %>%
  column_spec (2:ncol(maryland),border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "100px")
```

```{r}
maryland_cities <- crim_survey %>%
  subset(state %in% "md")

kable(maryland_cities) %>%
  column_spec (3:ncol(maryland_cities),border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "250px")
```

### Massachusetts 

```{r}
massachusetts <- type_by_state %>%
  filter(state == "ma") 

massachusetts <- massachusetts %>%
  mutate(total_laws = sum(massachusetts[,c(2:10)], na.rm=TRUE)) %>%
  select(1,11,2:10)
         
kable(massachusetts) %>%
  column_spec (2:ncol(massachusetts),border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "100px")
```

```{r}
massachusetts_cities <- crim_survey %>%
  subset(state %in% "ma")

kable(massachusetts_cities) %>%
  column_spec (3:ncol(massachusetts_cities),border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "300px")
```

### New York

```{r}
new_york <- crim_survey %>%
  subset(state %in% "ny")

kable(new_york) %>%
  column_spec (3:ncol(new_york),border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "175px")
```

### North Carolina

```{r}
north_carolina <- crim_survey %>%
  subset(state %in% "nc")

kable(north_carolina) %>%
  column_spec (3:ncol(north_carolina),border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "250px")
```

### Oregon

```{r}
oregon <- type_by_state %>%
  filter(state == "or") 

oregon <- oregon %>%
  mutate(total_laws = sum(oregon[,c(2:10)], na.rm=TRUE)) %>%
  select(1,11,2:10)

kable(oregon) %>%
  column_spec (2:ncol(oregon),border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "100px")
```

```{r}
oregon_cities <- crim_survey %>%
  subset(state %in% "or")

kable(oregon_cities) %>%
  column_spec (3:ncol(oregon_cities),border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "250px")
```

## Encampment Analysis

The following is an analysis of laws only related to homeless encampments: sleeping in public, camping in public, sitting/laying in public, and lodging/living/sleeping in vehicles or parking a vehicle used as a lodging.

### Encampment Law Overview

```{r}
encampment_totals <- law_totals %>%
  filter(key == "camping_in_public" |
           key == "sitting_laying_in_public" |
           key == "sleeping_in_public" |
           key == "living_in_vehicle")

kable(encampment_totals) %>%
  column_spec (2:ncol(encampment_totals),border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "250px")

encampment_type_totals <- law_type_totals %>%
  filter(key == "camping_in_public" |
           key == "sitting_laying_in_public" |
           key == "sleeping_in_public" |
           key == "living_in_vehicle")

kable(encampment_type_totals) %>%
  column_spec (2:ncol(encampment_type_totals),border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "250px")
```

### Encampment Laws by State and City

```{r}
encampments <- crim_survey_initial %>%
  ##get rid of columns we don't need to analyze
  filter(complete == "yes",
         coc != "state") %>%
  select(-name, -complete, -new_reporting_or_nlchp_fact_check, -c3, -meet_definition, -contains ("link"), -contains("law")) %>%
  ##reorder columns with mutated columns first
  select(3,1,2,4:7) 

encampments <- encampments %>%
  mutate(citywide = rowSums(encampments == "citywide", na.rm=TRUE),   
         particular_places = rowSums(encampments == "only in particular places", na.rm=TRUE),
         none = rowSums(encampments == "none", na.rm=TRUE),
         unclear = rowSums(encampments == "unclear", na.rm=TRUE))

encampments <- encampments %>%  
  mutate(total_laws = rowSums(encampments[,c("citywide", "particular_places")], na.rm=TRUE)) %>%
  select(1:3,12,8:11,4:7) %>%
  arrange(state)
  
kable(encampments) %>%
  column_spec (3:ncol(encampments),border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "400px")
```

### Encampment Laws by State
```{r}
encampment_state <- encampments %>%
  select(1,4:6,9:12) %>%
  group_by(state) %>%
  naniar::replace_with_na_all(condition = ~.x == "unclear") %>%
  naniar::replace_with_na_all(condition = ~.x == "none")

encampment_state[encampment_state == "citywide"] <- 1
encampment_state[encampment_state == "only in particular places"] <- 1
encampment_state[encampment_state == "county"] <- 1

encampment_state[,2:8] <- sapply(encampment_state[,2:8], as.numeric)
  encampment_state <- aggregate(encampment_state[,2:8], by=list(encampment_state$state), FUN=sum, na.rm=TRUE) %>%
  rename("state" = "Group.1")
  
kable(encampment_state) %>%
  column_spec (3:ncol(encampment_state),border_left = T, border_right = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "400px")
```

## Cross Examine CoCs with Homeless Counts

Read in HUD PIT counts and join to crim survey
```{r}
pit <- read_csv("~/GitHub/homelessness-in-the-us/data/output-data/pit-hic-zillow.csv") %>%
  rename(coc = coc_name)

homeless_crim <- left_join(crim_survey, pit, by="coc")
```

```{r, include=FALSE}
## Write csv files for adam with cocs as rows and Y/N laws as columns

type_by_coc_adam <- type_by_coc
type_by_coc_adam[type_by_coc_adam == "1"] <- "Y"
type_by_coc_adam[type_by_coc_adam == "0"] <- "N"
#view(type_by_coc_adam)

#write.csv(type_by_coc_adam, "crim-survey-adam.csv")
```

```{r, include=FALSE}
## Write csv files for all state summaries
#setwd("~/Documents/GitHub/homelessness-in-the-us/coc-crim-survey")
#write.csv(alaska, "alaska-summary.csv")
#write.csv(california, "california-summary.csv")
#write.csv(dc, "dc-summary.csv")
#write.csv(florida, "florida-summary.csv")
#write.csv(hawaii, "hawaii-summary.csv")
#write.csv(kentucky, "kentucky-summary.csv")
#write.csv(louisiana, "louisiana-summary.csv")
#write.csv(maryland, "maryland-summary.csv")
#write.csv(massachusetts, "massachusetts-summary.csv")
#write.csv(new_york, "new-york-summary.csv")
#write.csv(north_carolina, "north-carolina-summary.csv")
#write.csv(oregon, "oregon-summary.csv")
```
