---
title: "Untitled"
author: "Roxanne Ready"
output: 
  html_document:
    df_print: paged
    toc: TRUE
    toc_float: TRUE
    toc_depth: 3
---

```{r setup, include=FALSE}
# Setup Knitr to display code output by default but suppress messages
knitr::opts_chunk$set(echo = TRUE, paged.print = TRUE, message = FALSE, warning = FALSE, rows.print=10)
```

## Libraries

```{r results = 'hide'}

library(tidyverse) # For tidy data
library(dplyr)
library(rlang) # To simplify enqo() to {{}} when building functions
library(here) # For consistent file pathing
library(odbc) # For connecting to the database
library(RMySQL) # For connecting to the database
library(stringdist) # Dependancy for refinr; do not install from binary
library(refinr) # For address standardization
library(devtools)

# Presentation
library(knitr)
library(kableExtra)

```

```{r}

x <- read.csv(here("data", "input-data", "CoC_Cluster.csv")) %>%
  filter(`Cluster..`== 3) %>%
  select(CoC.Number, CoC.Name, `Cluster..`) %>%
  # Add col to designate state for grouping
  separate(CoC.Number, sep = "-", into = c("state_code", "coc_number"), remove = F) %>%
  # Drop the useless stand-alone number
  select(-coc_number) %>%
  select(state_code, CoC.Name, CoC.Number)

```


* Ratio of total homeless people (sheltered and unsheltered) to sheltered beds
* And ratio of total people to sheltered beds
* Grouped by CoCs
* Grouped by state


Reporter doing analysis on criminal cases
Login for case search -- access bulk data
Process for requesting bulk data

Okaloosa & flagler

IT department
772-288-5983

```{r}

x <- read.csv(here("data", "output-data", "pit-hic-zillow.csv")) %>%
  select(coc_code, state_code, coc_name, 
         coc_total_population_zillow_2017, 
         pit_overall_homeless_2017, 
         pit_sheltered_total_homeless_2017, pit_unsheltered_homeless_2017, 
         hic_total_year_round_beds_es_th_sh_2017)

wk <- x %>%
  # Total homeless people to sheltered beds
  mutate(homeless_to_beds = pit_overall_homeless_2017/hic_total_year_round_beds_es_th_sh_2017) %>%
  # Total population to sheltered beds
  mutate(population_to_beds = coc_total_population_zillow_2017/hic_total_year_round_beds_es_th_sh_2017) %>%
  group_by(state_code) %>%
  mutate(homeless_to_beds_state = sum(pit_overall_homeless_2017, na.rm = T)/sum(hic_total_year_round_beds_es_th_sh_2017, na.rm = T)) %>%
  mutate(population_to_beds_state = sum(coc_total_population_zillow_2017, na.rm = T)/sum(hic_total_year_round_beds_es_th_sh_2017, na.rm = T)) %>%
  left_join(read.csv(here("data", "input-data", "clean", "zillow-cluster.csv")) %>%
              select(coc_code, cluster_number_zillow_2017)) %>%
  filter(cluster_number_zillow_2017 == 3) %>%
  select(-cluster_number_zillow_2017) %>%
  select(coc_code, state_code, coc_name,
         homeless_to_beds, homeless_to_beds_state,
         population_to_beds, population_to_beds_state)

```

Sean L. Mussenden
202-590-2190
smussend@umd.edu
Knight Hall, 7765 Alumni Dr.
College Park, MD 20742

"Case Lists"

I'm looking for the following fields:
Case number
Address
Dates opened/entered & disposed
Offense and arrest dates
Charge and Statute
Case status
Agency and Jurisdiction
Disposition





Good afternoon, 

I'm a reporter with the Howard Center for Investigative Journalism and I'm hoping you can direct me to the proper method to request bulk court data access. Rather than providing a report, other counties in Florida have allowed us to use a log-in and access the data directly via their case-search systems. Either would be acceptable to us, provided the report is in a CSV or similar format, but I expect providing us access would be easier on your end.

We are seeking bulk access to all fields for criminal cases returned by date range for at least the previous 5 years. Fields should include but are not limited to:
Case number
Defendant Name
Defendant ID
Defendant Address
Agency 
Jurisdiction
Date Opened / Entered
Date Disposed
Offense Date
Arrest Date
Charge 
Statute
Case Status
Plea
Disposition / Verdict / Sentence
All fields related to Sentence, Probation, and Suspended times

I look forward to hearing back with next steps. Thank you very much for your time,


Good afternoon, 

I'm a reporter with the Howard Center for Investigative Journalism, requesting bulk court data access (please see attached for a list of the types of variables we are requesting access to).

Rather than providing a report, other counties in Florida have allowed us to use a log-in and access the data directly via their case-search systems. Either would be acceptable to us, provided the report is in a CSV or similar database-style format, but I expect providing us access would be easier on your end. We have filled out the Online Viewing Agreement, attached, to facilitate the process.

Please note we may need a customized access level to be able to view the data in bulk as requested. Additionally, if this kind of access is not available through your system, please advise as to next steps for requesting the bulk data.

Thank you very much for your time,


Roxanne Ready from the Howard Center for Investigative Journalism

You sent me an email yesterday regarding my bulk data request

If someone were to put in a request for the database itself, how would you handle that? We thought a login would be easier on your end, but since the database clearly exists for the website to draw from





















