---
title: "Preliminary Analysis"
author: "Theresa Diffendal, Sean Mussenden, Roxanne Ready"
date: "8/27/2019"
always_allow_html: yes
output: 
  md_document:
    variant: markdown_github
  html_document:
    toc: false
    df_print: paged
---

```{r setup, include=FALSE}
# Save this file and run the following line from the R console to output both HTML and .md formats:
# rmarkdown::render('documentation/memos/initial-analysis-memo.Rmd', output_format = 'all')

# Setup Knitr to display code output by default but suppress messages
knitr::opts_chunk$set(echo = TRUE, paged.print = TRUE, message = FALSE)

# Set the Knitr root directory to the project directory, which is one up from where this document is stored
# (Does not seem to affect links outside code blocks)
knitr::opts_knit$set(root.dir = '../..')
```

```{r include=FALSE}
#######################
#### Load Packages ####
#######################

library(tidyverse)
library(janitor)
library(readxl)
library(knitr)
library(kableExtra)

##############################
##### VARIABLES ##############
##############################

load_path <- paste0(here::here(), "/data/input-data/")

save_path <- paste0(here::here(), "/documentation/memos/output-csvs/")

##############################
##### LOAD DATA ##############
##############################

# Load and join Point in Time (PIT) homeless counts from HUD 2007-2018
hud_pit_14_18 <- read_xlsx(paste0(load_path, "2007-2018-PIT-Counts-by-CoC.xlsx"), sheet = 1, col_names=TRUE) %>% # 2018
  # left_join(read_xlsx("./data/input-data/2007-2018-PIT-Counts-by-CoC.xlsx", sheet = 2, col_names=TRUE), by="CoC Number") %>% # 2017
  # left_join(read_xlsx("./data/input-data/2007-2018-PIT-Counts-by-CoC.xlsx", sheet = 3, col_names=TRUE), by="CoC Number") %>% # 2016
  # left_join(read_xlsx("./data/input-data/2007-2018-PIT-Counts-by-CoC.xlsx", sheet = 4, col_names=TRUE), by="CoC Number") %>% # 2015
  left_join(read_xlsx(paste0(load_path, "2007-2018-PIT-Counts-by-CoC.xlsx"), sheet = 5, col_names=TRUE), by=c("CoC Number", "CoC Name")) %>% # 2014
  # left_join(read_xlsx("./data/input-data/2007-2018-PIT-Counts-by-CoC.xlsx", sheet = 6, col_names=TRUE), by="CoC Number") %>% # 2013
  # left_join(read_xlsx("./data/input-data/2007-2018-PIT-Counts-by-CoC.xlsx", sheet = 7, col_names=TRUE), by="CoC Number") %>% # 2012
  # left_join(read_xlsx("./data/input-data/2007-2018-PIT-Counts-by-CoC.xlsx", sheet = 8, col_names=TRUE), by="CoC Number") %>% # 2011
  # left_join(read_xlsx("./data/input-data/2007-2018-PIT-Counts-by-CoC.xlsx", sheet = 9, col_names=TRUE), by="CoC Number") %>% # 2010
  # left_join(read_xlsx("./data/input-data/2007-2018-PIT-Counts-by-CoC.xlsx", sheet = 10, col_names=TRUE), by="CoC Number") %>% # 2009
  # left_join(read_xlsx("./data/input-data/2007-2018-PIT-Counts-by-CoC.xlsx", sheet = 11, col_names=TRUE), by="CoC Number") %>% # 2008
  # left_join(read_xlsx("./data/input-data/2007-2018-PIT-Counts-by-CoC.xlsx", sheet = 12, col_names=TRUE), by="CoC Number") %>% # 2007
  # Clean up column names
  janitor::clean_names() %>%
  rename_all(~stringr::str_replace_all(., "co_c", "coc")) %>%
  rename(coc_code = coc_number) %>%
  # Clean up stray "."s
  naniar::replace_with_na_all(condition = ~.x == ".") %>%
  # Convert all numeric cols to numeric
  mutate_at(vars(4:125), as.numeric) %>%
  # Standardize all characters to lowercase
  mutate_if(is.character, tolower) %>%
  # Add col to designate state
  separate(coc_code, sep = "-", into = c("state_code", "coc_number"), remove = F) 

# Load Zillow's cluster analysis file from 2017
zillow_cluster <- read_csv("./data/input-data/CoC_Cluster.csv") %>%
  # Clean up column names
  janitor::clean_names() %>%
  rename_all(~stringr::str_replace_all(., "co_c", "coc")) %>%
  rename(coc_code = coc_number) %>%
  # Clean up stray "."s
  naniar::replace_with_na_all(condition = ~.x == ".") %>%
  # Convert all numeric cols to numeric
  mutate_at(vars(3:8), as.numeric) %>%
  # Standardize all characters to lowercase
  mutate_if(is.character, tolower) %>%
  # Append year to numeric cols for clarity
  dplyr::rename_if(is.numeric, function(x) paste0(x, "_cluster_2017")) %>%
  # Add col to designate state
  separate(coc_code, sep = "-", into = c("state_code", "coc_number"), remove = F) 

# Join cluster analyses of 2017 with HUD PIT data for rate estimates
hud_pit_zillow_cluster <- hud_pit_14_18 %>%
  full_join(zillow_cluster, by= c("coc_code", "state_code", "coc_number")) %>%
  select(state_code, coc_code, coc_name = coc_name.x, overall_homeless_2014, overall_homeless_2018, coc_total_population_cluster_2017, estimated_homeless_rate_percent_cluster_2017) %>%
  mutate(# Change in overall homelessness 
         overall_change_2014_2018 = overall_homeless_2018 -  overall_homeless_2014,
         overall_pct_change_2014_2018 = (overall_homeless_2018 -  overall_homeless_2014)/overall_homeless_2014,
         # Rates of homelessness for 2014 and 2018 based on available 2017 population numbers
         homeless_rate_2014 = overall_homeless_2014/coc_total_population_zillow_2017*100,
         homeless_rate_2018 = overall_homeless_2018/coc_total_population_zillow_2017*100,
         # Change in homelessness rates
         overall_pct_pt_change_2014_2018 = homeless_rate_2018 - homeless_rate_2014
         )

```

## Introduction

### Purpose of this memo

This memo includes code and results from a preliminary exploration of the data for use by field reporters in narrowing their initial reporting. Accompanying this memo are four .csv files for use by the reporters. They can be found [here](https://github.com/shardsofblue/homelessness-project-fall2019/tree/master/documentation/memos/output-csvs) (Click the file, right-click "Raw", choose "Save Link As", and change the file extention from ".txt" to ".csv").

### Definitions

* **Continuum of Care**: "A Continuum of Care (CoC) is a regional or local planning body that coordinates housing and services funding for homeless families and individuals," (The 2018 Annual Homeless Assessment Report to Congress p. 2). CoCs are not consistent in size, shape or population.

### Important caveats

**Findings herein are preliminary estimates and should not be published.** Please bear the following caveats in mind when reviewing this memo:

1. Methods of counting the population of people experiencing homelessness may have changed over the years, which may have caused flucutations in counts that do not reflect actual changes.
2. Because we do not have estimates for the total population of CoCs for 2014 and 2018 yet, we used an available 2017 population estimate to calculate rates of change in the homeless population at 2014 and 2018. These rates are therefore only rough estimations, to be perfected when we have more accurate overall population data.
3. D.C. is a consistent outlier in statewide groupings because, unlike most other state groups, it includes only one city, and cities generally have higher rates of homelessness that rural areas.
4. Some Continuums of Care have changed over the years, merging together or into new CoCs. The following CoCs have known issues during the period covered here, having merged with other CoCs: CT-502, CT-506, CT-503, CT-508, CT-512, CT-505, FL-516, FL-503, MA-513, MA-516, MA-520, MA-511, NJ-518, NJ-503, NY-509, NY-502, NY-505, NY-517, 	NY-508, PA-507, PA-509, PA-602, PA-601, TX-703, TX-607, IN-500, IN-502, LA-504, LA-509, MA-518, MA-516, ME-502, ME-500, AR-512, AR-503, LA-508, LA-509, WA-507, WA-501.

## Findings

### Fact: Highest rates of homelessness by Continuum of Care 2017

To find the rate of homelessness for an area, we divided the count of people experiencing homelessness for a given year by the total population in 2017 (the only year for which we have this data at the CoC level). 

In 2017, Mendocino County's CoC had the nation's highest rate of homelessness, at 2 percent, followed by Washington, D.C. (1.2 percent).  

Five of the top 10 were in California (Mendocino County, Santa Cruz area, San Francisco, Monterey area, Imperial County).  Other areas with high rates: Boston, New York City, Atlanta and the Florida Keys (Monroe County).

#### Supporting code and output

**Rates of Homelessness by CoC, Highest to Lowest**
```{r, rows.print=20}

# CoCs by estimated homeless percent 2017, highest to lowest
zillow_cluster %>%
  select(state_code, coc_code, estimated_homeless_rate_percent_cluster_2017, coc_total_population_cluster_2017, coc_name) %>%
  arrange(desc(estimated_homeless_rate_percent_cluster_2017)) %>%
  mutate(rank_rate = rank(desc(estimated_homeless_rate_percent_cluster_2017), ties.method= "first")) %>%
  write_csv(paste0(save_path, "rates-of-homelessness-by-coc.csv")) %>%
  kable(caption = "Rates of Homelessness by CoC, Highest to Lowest") %>%
  kable_styling("striped", fixed_thead = T) %>%
  scroll_box(width = "100%", height = "500px")

```


**Rates of Homelessness by CoC, Lowest to Highest**
```{r, rows.print=20}

# CoCs by estimated homeless percent 2017, lowest to highest
zillow_cluster %>%
  select(state_code, coc_code, estimated_homeless_rate_percent_cluster_2017, coc_total_population_cluster_2017, coc_name) %>%
  arrange(estimated_homeless_rate_percent_cluster_2017) %>%
  mutate(rank_rate = rank(desc(estimated_homeless_rate_percent_cluster_2017), ties.method= "first")) %>%
  kable() %>%
  kable_styling("striped", fixed_thead = T) %>%
  scroll_box(width = "100%", height = "500px")

```

**Counts of CoCs in the Top 10**
```{r}
# How many of the top 10 were from each state?
zillow_cluster %>%
  select(state_code, coc_code, est_rate_perc_2017 = estimated_homeless_rate_percent_cluster_2017) %>%
  arrange(desc(est_rate_perc_2017)) %>%
  top_n(10) %>%
  select(state_code) %>%
  group_by(state_code) %>%
  summarize(count_coc = n()) %>%
  arrange(desc(count_coc)) %>%
  kable() %>%
  kable_styling("striped", fixed_thead = T) %>%
  scroll_box(width = "100%", height = "350px")

```

### Fact: Highest rates of homelessness by state

To generate an overview by state, we averaged homelessness rates in all CoCs within a state together. _See Caveat #2 above._

Washington, D.C., followed by Hawaii, California and Oregon had highest average rates of homelessness. Illinois, Louisiana and Mississippi had the lowest. 

#### Supporting code and output

**Rates of Homelessness by State, Highest to Lowest**
```{r, rows.print=20}

# Homelessness rates by state 2017, highest to lowest
zillow_cluster %>%
  select(state_code, coc_name, coc_total_population_cluster_2017, est_rate_perc_2017 = estimated_homeless_rate_percent_cluster_2017) %>%
  na.omit() %>%
  group_by(state_code) %>%
  summarise(mean_homeless_rate = mean(est_rate_perc_2017),
            total_pop_2017 = sum(coc_total_population_cluster_2017)) %>%
  arrange(desc(mean_homeless_rate)) %>%
  mutate(rank_rate = rank(desc(mean_homeless_rate), ties.method= "first")) %>%
  select(rank_rate, everything()) %>%
  write_csv(paste0(save_path, "rates-of-homelessness-by-state.csv")) %>%
  kable() %>%
  kable_styling("striped", fixed_thead = T) %>%
  scroll_box(width = "100%", height = "500px")

```

**Rates of Homelessness by State, Lowest to Highest**
```{r, rows.print=20}

# Homelessness rates by state 2017, highest to lowest
zillow_cluster %>%
  select(state_code, coc_name, coc_total_population_cluster_2017, est_rate_perc_2017 = estimated_homeless_rate_percent_cluster_2017, coc_total_population_cluster_2017) %>%
  na.omit() %>%
  group_by(state_code) %>%
  summarise(mean_homeless_rate = mean(est_rate_perc_2017),
            total_pop_2017 = sum(coc_total_population_cluster_2017)) %>%
  arrange(mean_homeless_rate) %>%
  mutate(rank_rate = rank(desc(mean_homeless_rate), ties.method= "first")) %>%
  select(rank_rate, everything()) %>%
  kable() %>%
  kable_styling("striped", fixed_thead = T) %>%
  scroll_box(width = "100%", height = "500px")

```

### Fact: Percent increases/decreases per CoC ROUGH ESTIMATE

_See Caveat #2 above._

We calculated the rates of homelessness in 2014 and 2018 for CoCs compared to total population estimates from 2017. This rough calculation shows that by percentage, Imperial County, CA; El Dorado County, CA; Lynn, Massachusetts and Norman/Cleveland County, Oklahoma had the highest overall increases, while tuscaloosa city & county, AL; elizabeth/union county, NJ; Fort Walton Beach/Okaloosa, Walton Counties, FL and Punta Gorda/Charlotte County, FL had the greatest decreases. _These are only rough estimates for use in preliminary reporting and should not be used in final stories._


#### Supporting code and output

**Percent Homeless Increase by CoC, Highest to Lowest**
```{r, rows.print = 20}
# Calculate ROUGH percent change by CoC
hud_pit_14_18 %>%
  left_join(zillow_cluster, by= c("coc_code", "state_code", "coc_number", "coc_name")) %>%
  select(state_code, coc_code, coc_name, overall_homeless_2014, overall_homeless_2018, coc_total_population_cluster_2017, est_rate_perc_2017 = estimated_homeless_rate_percent_cluster_2017) %>%
  mutate(homeless_rate_2014 = overall_homeless_2014/coc_total_population_cluster_2017*100,
         homeless_rate_2018 = overall_homeless_2018/coc_total_population_cluster_2017*100,
         overall_pct_change_2014_2018 = (overall_homeless_2018 -  overall_homeless_2014)/overall_homeless_2014
         ) %>%
  na.omit() %>%
  arrange(desc(overall_pct_change_2014_2018)) %>%
  select(state_code, coc_code, coc_name, overall_pct_change_2014_2018, homeless_rate_2014, homeless_rate_2018, coc_total_population_cluster_2017) %>%
  write_csv(paste0(save_path, "homelessness-increases-by-coc.csv")) %>%
  kable() %>%
  kable_styling("striped", fixed_thead = T) %>%
  scroll_box(width = "100%", height = "500px")
```

**Percent Homeless Increase by CoC, Lowest to Highest**
```{r, rows.print = 20}
# Calculate ROUGH percent change by CoC
hud_pit_14_18 %>%
  left_join(zillow_cluster, by= c("coc_code", "state_code", "coc_number", "coc_name")) %>%
  select(state_code, coc_code, coc_name, overall_homeless_2014, overall_homeless_2018, coc_total_population_cluster_2017, est_rate_perc_2017 = estimated_homeless_rate_percent_cluster_2017, coc_total_population_cluster_2017) %>%
  mutate(homeless_rate_2014 = overall_homeless_2014/coc_total_population_cluster_2017*100,
         homeless_rate_2018 = overall_homeless_2018/coc_total_population_cluster_2017*100,
         overall_pct_change_2014_2018 = (overall_homeless_2018 -  overall_homeless_2014)/overall_homeless_2014,
         ) %>%
  na.omit() %>%
  arrange(overall_pct_change_2014_2018) %>%
  select(state_code, coc_code, coc_name, overall_pct_change_2014_2018, homeless_rate_2014, homeless_rate_2018, coc_total_population_cluster_2017) %>%
  kable() %>%
  kable_styling("striped", fixed_thead = T) %>%
  scroll_box(width = "100%", height = "500px")
```

**Counts of CoCs in the Top 10 Greatest Increases**
```{r, rows.print = 20}
# How many were from each state?
hud_pit_14_18 %>%
  # Top 10
  left_join(zillow_cluster, by= c("coc_code", "state_code", "coc_number", "coc_name")) %>%
  select(state_code, coc_name, overall_homeless_2014, overall_homeless_2018, coc_total_population_cluster_2017, est_rate_perc_2017 = estimated_homeless_rate_percent_cluster_2017) %>%
  # Compare homeless counts at given years to total population in 2017
  mutate(homeless_rate_2014 = overall_homeless_2014/coc_total_population_cluster_2017*100,
         homeless_rate_2018 = overall_homeless_2018/coc_total_population_cluster_2017*100,
          # Change in overall homelessness
         overall_pct_change_2014_2018 = (overall_homeless_2018 - overall_homeless_2014)/overall_homeless_2014
         ) %>%
  na.omit() %>%
  arrange(desc(overall_pct_change_2014_2018)) %>%
  mutate(rank = rank(desc(overall_pct_change_2014_2018))) %>%
  filter(between(rank, 1, 10)) %>%
  group_by(state_code) %>%
  summarize(count_coc_top = n()) %>%
  select(state_code, count_coc_top) %>%
  # Bottom 10
  full_join(
    hud_pit_14_18 %>%
    left_join(zillow_cluster, by= c("coc_code", "state_code", "coc_number", "coc_name")) %>%
    select(state_code, coc_name, overall_homeless_2014, overall_homeless_2018, coc_total_population_cluster_2017, est_rate_perc_2017 = estimated_homeless_rate_percent_cluster_2017) %>%
    mutate(homeless_rate_2014 = overall_homeless_2014/coc_total_population_cluster_2017*100,
           homeless_rate_2018 = overall_homeless_2018/coc_total_population_cluster_2017*100,
           overall_pct_change_2014_2018 = (overall_homeless_2018 - overall_homeless_2014)/overall_homeless_2014
           ) %>%
    na.omit() %>%
    arrange(desc(overall_pct_change_2014_2018)) %>%
    mutate(rank = rank(overall_pct_change_2014_2018)) %>%
    filter(between(rank, 1, 10)) %>%
    group_by(state_code) %>%
    summarize(count_coc_bottom = n()) %>%
    select(state_code, count_coc_bottom) 
  ) %>%
  kable() %>%
  kable_styling("striped", fixed_thead = T) %>%
  scroll_box(width = "100%", height = "500px")

```

### Fact: Percent increases/decreases per state ROUGH ESTIMATE

_See Caveat #2 above._

We calculated the rates of homelessness in 2014 and 2018 compared to total population estimates from 2017. This rough calculation shows that by percentage, South Dakota, Colorado and Delaware had the highest overall increases, while North Dakota, Mississippi and Kentucky had the greatest decreases. _These are only rough estimates for use in preliminary reporting and should not be used in final stories._

#### Supporting code and output

**Percent Homeless Increase by State, Highest to Lowest**
```{r, rows.print=20}

# Calculate ROUGH percent change by state
hud_pit_zillow_cluster %>%
  select(state_code, coc_name, 
         overall_homeless_2014, overall_homeless_2018, 
         coc_total_population_cluster_2017, 
         overall_pct_change_2014_2018, 
         est_rate_perc_2017 = estimated_homeless_rate_percent_cluster_2017) %>%
  na.omit() %>%
  # Group by states
  group_by(state_code) %>%
  # Average the changes for statewide average change
  summarize(
    mean_overall_homeless_2014 = mean(overall_homeless_2014),
    mean_overall_homeless_2018 = mean(overall_homeless_2018),
    mean_overall_pct_change_2014_2018 = mean(overall_pct_change_2014_2018),
    total_pop_2017 = sum(coc_total_population_cluster_2017)
  ) %>%
  #mutate(test = (mean_homeless_rate_2018 - mean_homeless_rate_2014) / mean_homeless_rate_2014) %>%
  arrange(desc(mean_overall_pct_change_2014_2018)) %>%
  mutate(rank = rank(desc(mean_overall_pct_change_2014_2018), ties.method= "first")) %>%
  select(rank, state_code, mean_overall_pct_change_2014_2018, mean_overall_homeless_2014, mean_overall_homeless_2018, total_pop_2017) %>%
  write_csv(paste0(save_path, "homelessness-increases-by-state.csv")) %>%
  kable() %>%
  kable_styling("striped", fixed_thead = T) %>%
  scroll_box(width = "100%", height = "500px")

```

**Percent Homeless Increase by State, Lowest to Highest**
```{r, rows.print=20}

# Calculate ROUGH percent change by state
hud_pit_zillow_cluster %>%
  select(state_code, coc_name, 
         overall_homeless_2014, overall_homeless_2018, 
         coc_total_population_cluster_2017, 
         overall_pct_change_2014_2018, 
         est_rate_perc_2017 = estimated_homeless_rate_percent_cluster_2017) %>%
  na.omit() %>%
  # Group by states
  group_by(state_code) %>%
  # Average the changes for statewide average change
  summarize(
    mean_overall_homeless_2014 = mean(overall_homeless_2014),
    mean_overall_homeless_2018 = mean(overall_homeless_2018),
    mean_overall_pct_change_2014_2018 = mean(overall_pct_change_2014_2018),
    total_pop_2017 = sum(coc_total_population_cluster_2017)
  ) %>%
  #mutate(test = (mean_homeless_rate_2018 - mean_homeless_rate_2014) / mean_homeless_rate_2014) %>%
  arrange(mean_overall_pct_change_2014_2018) %>%
  mutate(rank = rank(desc(mean_overall_pct_change_2014_2018), ties.method= "first")) %>%
  select(rank, state_code, mean_overall_pct_change_2014_2018, mean_overall_homeless_2014, mean_overall_homeless_2018, total_pop_2017) %>%
  kable() %>%
  kable_styling("striped", fixed_thead = T) %>%
  scroll_box(width = "100%", height = "500px")

```










