---
title: "dc encampments"
author: "Sean Mussenden | Howard Center for Investigative Journalism"
date: "2/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Packages

```{r}
library(tidyverse)
library(janitor)
library(tidycensus)
library(sf)
library(leaflet)
library(tigris)
library(readxl)

census_api_key("549950d36c22ff16455fe196bbbd01d63cfbe6cf")

```

### Reading in the data
```{r}

# 
# https://docs.google.com/spreadsheets/d/1Jsh_KUi9Ww5pa2V5swXNx_2G2P7Nv6mf8q7JVJrEJPU/edit

# Read in encampment locations with lat long, remove for moment those locations with no geocoordinates, create an empty column to hold geocoded census tract. 
encampment_locations <- read_csv("data/updated_data/encamplent_locations__2015_2019.csv") %>%
  clean_names() %>%
  filter(!is.na(coordinates)) %>%
  mutate(census_block = "") 

# Geocode locations to census tract.  Really can't figure out tidy way to do this, so this is what we're doing for now. 
encampment_locations$census_block <- apply(encampment_locations, 1, function(row) call_geolocator_latlon(row['latitude'], row['longitude']))

# Read in encampment cleaings (one record per clearing, with datetime)
encampment_clearings <- read_csv("data/updated_data/dc_encampment_clearings_2015_2019_ cleaned_data.csv") %>%
  select(-notes, -date_full)

# Join them together
encampment_clearings_locations <- encampment_clearings %>%
  inner_join(encampment_locations, by=c("location")) 

# Write it out so we don't have to geocode all over again. 
write_csv(encampment_clearings_locations, "data/updated_data/geolocated_clearances.csv")

# Clean up everything
rm(list = ls())

# Read in the csv we just created
encampment_clearings_locations <- read_csv("data/updated_data/geolocated_clearances.csv") %>%
  mutate_all(as.character)

# Convert census blocks columns to census tracts by taking only first 11 characters of 14
encampment_clearings_locations <- encampment_clearings_locations %>%
  mutate(census_tract = str_sub(census_block, 1, 11)) %>%
  select(-census_block)

# Read in gentrification data
gentrification <- read_xlsx("data/updated_data/US_tr_GentDecline.xlsx") %>%
  clean_names() %>%
  filter(state_name == "District of Columbia") %>%
  select(geoid, strong_expansion_decline, tot_pop16) %>%
  rename(low_income_loss_gain = strong_expansion_decline) %>%
  mutate(low_income_pct_loss_gain = low_income_loss_gain/tot_pop16)

# Join it to clearances_data

encampment_clearings_locations <- encampment_clearings_locations %>%
  inner_join(gentrification, by=c("census_tract" = "geoid"))

# Get median income by census tract in DC in 2018 and 2011, and then join them together and compute income change over that 8 year period

dc_income_2018 <- get_acs(state = "DC", geography = "tract", 
                  variables = "B06011_001", year = 2018) %>%
  mutate(GEOID = as.character(GEOID)) %>%
  rename(median_household_income_2018 = estimate) %>% 
  select(-variable, -moe)

dc_income_2011 <- get_acs(state = "DC", geography = "tract", 
                  variables = "B06011_001", year = 2011) %>%
  mutate(GEOID = as.character(GEOID)) %>%
  rename(median_household_income_2011 = estimate) %>% 
  select(-variable, -moe)

dc_income_change <- dc_income_2011 %>%
  inner_join(dc_income_2018) %>%
  clean_names() %>%
  select(-name) %>%
  mutate(income_change_2011_2018 = round((median_household_income_2018-median_household_income_2011)/median_household_income_2011,2)) %>%
  select(-median_household_income_2018, -median_household_income_2011)

# Join it to clearances data

encampment_clearings_locations <- encampment_clearings_locations %>%
  inner_join(dc_income_change, by=c("census_tract" = "geoid"))

# Get population by census tract in DC in 2018 and 2011, and then join them together and compute population total change and percent change
#acs_variables <- load_variables(2018, "acs5", cache = TRUE)

dc_population_2018 <- get_acs(state = "DC", geography = "tract", 
                  variables = "B01001_001", year = 2018) %>%
  mutate(GEOID = as.character(GEOID)) %>% 
  rename(total_population_2018 = estimate) %>% 
  select(-variable, -moe) 

dc_population_2011 <- get_acs(state = "DC", geography = "tract", 
                  variables = "B01001_001", year = 2011) %>%
  mutate(GEOID = as.character(GEOID)) %>% 
  rename(total_population_2011 = estimate) %>% 
  select(-variable, -moe) 

dc_population_change <- dc_population_2011 %>%
  inner_join(dc_population_2018) %>%
  clean_names() %>%
  select(-name) %>%
  mutate(population_change_2011_2018_pct = round((total_population_2018-total_population_2011)/total_population_2011,2)) %>%
  mutate(total_population_change_2011_2018 = total_population_2018-total_population_2011) %>%
  select(-total_population_2018, -total_population_2011)

# Join it to clearances data

encampment_clearings_locations <- encampment_clearings_locations %>%
  inner_join(dc_population_change, by=c("census_tract" = "geoid"))

# Group it by census tract 

clearings_summary <- encampment_clearings_locations %>%
  group_by(census_tract) %>%
  summarise(clearances = n(),
    low_income_pct_loss_gain = mean(low_income_pct_loss_gain),
    low_income_loss_gain = mean(low_income_loss_gain),
    income_change_2011_2018 = mean(income_change_2011_2018),
    population_change_2011_2018_pct = mean(population_change_2011_2018_pct),
    total_population_change_2011_2018 = mean(total_population_change_2011_2018)
            ) %>%
  arrange(desc(clearances))

write_csv(clearings_summary, "data/updated_data/clearings_summary.csv")



```

# Remove non-coordinate locations

dc_encampments_sf <- dc_encampments %>%
  st_as_sf(coords=c("longitude","latitude"), crs=4269, remove=FALSE)

dc_encampments_plus_tracts <- read_csv("data/encampments_dc_plus_tracts.csv") %>%
  mutate(GEOID = as.character(GEOID))

Summary(dc_encampments_plus_tracts)

print(encam)

```
```{r}
dc_income_2018 <- get_acs(state = "DC", geography = "tract", 
                  variables = "B06011_001", year = 2018) %>%
  mutate(GEOID = as.character(GEOID)) %>%
  rename(median_household_income_2018 = estimate) %>% 
  select(-variable, -moe)

dc_income_2017 <- get_acs(state = "DC", geography = "tract", 
                  variables = "B06011_001", year = 2017) %>%
  mutate(GEOID = as.character(GEOID)) %>%
  rename(median_household_income_2017 = estimate) %>% 
  select(-variable, -moe)

dc_income_2016 <- get_acs(state = "DC", geography = "tract", 
                  variables = "B06011_001", year = 2016) %>%
  mutate(GEOID = as.character(GEOID)) %>%
  rename(median_household_income_2016 = estimate) %>% 
  select(-variable, -moe)

dc_income_2015 <- get_acs(state = "DC", geography = "tract", 
                  variables = "B06011_001", year = 2015) %>%
  mutate(GEOID = as.character(GEOID)) %>%
  rename(median_household_income_2015 = estimate) %>% 
  select(-variable, -moe)

dc_income_2014 <- get_acs(state = "DC", geography = "tract", 
                  variables = "B06011_001", year = 2014) %>%
  mutate(GEOID = as.character(GEOID)) %>%
  rename(median_household_income_2014 = estimate) %>% 
  select(-variable, -moe)

dc_income_2013 <- get_acs(state = "DC", geography = "tract", 
                  variables = "B06011_001", year = 2013) %>%
  mutate(GEOID = as.character(GEOID)) %>%
  rename(median_household_income_2013 = estimate) %>% 
  select(-variable, -moe)

dc_income_2012 <- get_acs(state = "DC", geography = "tract", 
                  variables = "B06011_001", year = 2012) %>%
  mutate(GEOID = as.character(GEOID)) %>%
  rename(median_household_income_2012 = estimate) %>% 
  select(-variable, -moe)

dc_income_2011 <- get_acs(state = "DC", geography = "tract", 
                  variables = "B06011_001", year = 2011) %>%
  mutate(GEOID = as.character(GEOID)) %>%
  rename(median_household_income_2011 = estimate) %>% 
  select(-variable, -moe)

dc_income_2010 <- get_acs(state = "DC", geography = "tract", 
                  variables = "B06011_001", year = 2010) %>%
  mutate(GEOID = as.character(GEOID)) %>%
  rename(median_household_income_2010 = estimate) %>% 
  select(-variable, -moe)

dc_income <- dc_income_2018 %>%
  inner_join(dc_income_2017, by=c("GEOID","NAME")) %>%
  inner_join(dc_income_2016, by=c("GEOID","NAME")) %>%
  inner_join(dc_income_2015, by=c("GEOID","NAME")) %>%
  inner_join(dc_income_2014, by=c("GEOID","NAME")) %>%
  inner_join(dc_income_2013, by=c("GEOID","NAME")) %>%
  inner_join(dc_income_2012, by=c("GEOID","NAME")) %>%
  inner_join(dc_income_2011, by=c("GEOID","NAME")) %>%
  inner_join(dc_income_2010, by=c("GEOID","NAME")) %>%
  mutate(chg_pct_2010_2018 = (median_household_income_2018-median_household_income_2010)/median_household_income_2010*100)

```


```{r}

dc_encampments_plus_tracts_income <- dc_encampments_plus_tracts %>%
  left_join(dc_income, by=c("GEOID"))
  
write.csv(x, "data/dc_encampments_plus_tracts_income.csv")


```


```{r}

# Income in census tracts most recent

x <- dc_encampments_plus_tracts_income %>%
  group_by(NAMELSAD) %>%
  summarise(count = n(),
            median_household_income_2018 = mean(median_household_income_2018),
            median_household_income_2010 = mean(median_household_income_2010),
            chg_pct_2010_2018 = mean(chg_pct_2010_2018)
            ) %>%
  arrange(desc(chg_pct_2010_2018))

write.csv(x, "data/dc_encampments_plus_tracts_income_summary_stats.csv")

# Income multiple years



```


```{r}
leaflet() %>%
  addTiles() %>%
  addMarkers(lng=dc_encampments$longitude, lat=dc_encampments$latitude) %>%
  addPolygons()
```

```{r}
options(tigris_use_cache = TRUE)


acs_variables <- load_variables(2017, "acs5", cache = TRUE)

# Get shapefiles with median income in past 12 months
dc_income <- get_acs(state = "DC", geography = "tract", 
                  variables = "B06011_001") %>%
  mutate(GEOID = as.character(GEOID))


```
```{r}

plot(dc_tracts$geometry)
plot(dc_encampments_sf$geometry, add=T, pch=16, cex=0.5, col="gray50")

```


```{r}
st_join(dc_tracts,dc_encampments_sf) 

```

a
a
df


```{r}

st_crs(dc_encampments_sf)
st_crs(dc_tracts)


```
